---
title: "Basic ALFA+ dataset"
author: "Clazz"
date: "2025-05-27"
output: html_document
---


```{r}
library(readxl)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("psych")
library(psych)
#install.packages("tidyverse")
library(tidyverse)
#install.packages("ggpubr")
library(ggpubr)
#install.packages("broom")
library(broom)
#install.packages("ggeffects")
library(ggeffects)
#install.packages("MASS")
library(MASS)
#install.packages("mice")
library(mice)
#install.packages("VIM")
library(VIM)
#install.packages("car")
library(car)
#install.packages("sjPlot")
library(sjPlot)
#install.packages("caret")
library(caret)
#install.packages("bestNormalize")
library(bestNormalize)
#install.packages("ggrain")
library(ggrain)
#install.packages("emmeans")
library(emmeans)
#install.packages("skimr")
library(skimr)
library(glmnet)
#install.packages("bullseye")
library(bullseye)
#install.packages("lavaan")
library(lavaan)
#install.packages("brms")
library(brms)
#install.packages("mgcv")
library(mgcv)
library(splines)
#install.packages("fitdistrplus")
library(fitdistrplus)
#install.packages("patchwork")
library(patchwork)
```

This code processes ALFA+ cohort data into a main dataset. variable specs can be found in the READ_ME file


#making the dataset

##alfa visits basic dataset

```{r}

master_nd<- read_excel("G:/My Drive/Data/ALFA+/new participants_V1_03042025.xlsx")%>%
  dplyr::select(ID, sex, age_v1,education)%>%
  rename(subject = ID)%>%
  mutate(subject = factor(subject),
         sex= factor(sex),
         age_v1 = as.numeric(age_v1))
bd<-read_excel("G:/My Drive/Data/ALFA+/20210517_APlus_General_Clinical.xlsx")%>%
  rename(subject = IdParticipante)%>%
   mutate(subject = factor(subject))%>%
  dplyr::select(subject, DOB)%>%
  unique()

master_nd<-left_join(master_nd, bd)

```



##TIV
```{r}
TIV<- read_excel("G:/My Drive/Data/ALFA+/TIV_ALFA+_V1.xlsx")

TIV<-TIV%>%
  dplyr::select(subject, value)%>%
  mutate(subject = factor(subject))%>%
  rename(TIV = value)

#add TIV
master_nd<-left_join(master_nd, TIV)

```

##read in the modifiable risk factors 

###women's reproductive history
I want to find out which of the women had more than 4 kids that they changed...
```{r}

#removing women who reported pregnancy related diseases 

women_preg<-read_delim("G:/My Drive/Data/ALFA+/reproductive_var/Alfa Plus Long 2025_1054_AF_CRF_AF_embarazos_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  filter(Visita == "V2")%>%
  dplyr::select(IdParticipante, AF_deliverypat, AF_deliveryage)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))%>%
  group_by(subject)%>%
  mutate(AF_deliverypat = max(AF_deliverypat),
         AF_deliveryage = min(AF_deliveryage))%>%
  ungroup()%>%
  unique()%>%
  rename(Age_at_first_birth = AF_deliveryage)

women_preg%>%
  filter(AF_deliverypat !=0)

master_nd<-left_join(master_nd, women_preg)

master_nd<-master_nd%>%
  mutate(AF_deliverypat = replace_na(AF_deliverypat, 0))%>%
  filter(AF_deliverypat == 0)%>%
  dplyr::select(-AF_deliverypat)



#age menopause + abortions + menarche 

menopause<-read_delim("G:/My Drive/Data/ALFA+/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, F06_2_017, F06_2_019, F06_2_018, F06_2_015)%>%
  rename(subject = IdParticipante,
         age_menopause = F06_2_017, 
         abortions = F06_2_019,
         preg_termino = F06_2_018,
         age_menarche = F06_2_015)%>%
  filter(!if_all(age_menopause:age_menarche, is.na))%>%
  group_by(subject)%>%
  mutate(age_menopause = max(age_menopause),
         abortions = max(abortions),
         age_menarche = min(age_menarche))%>%
  ungroup()%>%
  unique()%>%
  mutate(subject = factor(subject),
         age_menopause = ifelse(age_menopause == 0, NA, age_menopause))

master_nd<-left_join(master_nd, menopause)

cause_menopause<-read_delim("G:/My Drive/Data/ALFA+/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, AF_menopcs)%>%
  rename(cause_menopause = AF_menopcs)%>%
  mutate(cause_menopause = ifelse(cause_menopause == -77 | cause_menopause == -66, NA, cause_menopause))%>%
  drop_na(cause_menopause)%>%
  mutate(cause_menopause = factor(ifelse(cause_menopause == 3, "non-natural", "natural")))%>%
  unique()%>%
  filter(!(IdParticipante == "11397" & cause_menopause == "natural"))%>%
  filter(!(IdParticipante == "55036" & cause_menopause == "natural"))%>%
  filter(!(IdParticipante == "11882" & cause_menopause == "natural"))#found duplicate entries becuse participants answers change

cause_early_menopause<-read_delim("G:/My Drive/Data/ALFA+/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, AF_earlymenopcs, AF_horm_sust)%>%
  mutate(AF_earlymenopcs = factor(AF_earlymenopcs))
  
cause_early_menopause%>%
  filter(AF_earlymenopcs == 2)%>%
  unique()%>%
  pull(IdParticipante)

cause_menopause$cause_menopause[cause_menopause$IdParticipante == 10225]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 10388]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 11111]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 11114]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 11557]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 44193]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 55036]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 55274]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 66301]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 77191]<-"non-natural"

cause_menopause<-cause_menopause%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))

menopause%>%
  drop_na(age_menopause)

master_nd<-left_join(master_nd, cause_menopause)

# HRT
hrt<-read_delim("G:/My Drive/Data/ALFA+/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, AF_horm_sust, Visita)%>%
  mutate(AF_horm_sust = factor(AF_horm_sust),
         IdParticipante = factor(IdParticipante))%>%
  rename(HRT = AF_horm_sust,
         subject = IdParticipante)%>%
  dplyr::select(-Visita)%>%
  unique()%>%
  na.omit(HRT)%>%
  group_by(subject)%>%
  filter(!(n() > 1 & HRT == 0)) %>%
  ungroup()

master_nd<-left_join(master_nd, hrt)

#replacing the no answers values with NA
master_nd<-master_nd%>%
  mutate(age_menopause = ifelse(age_menopause == -77, NA, age_menopause),
         age_menarche = ifelse(age_menarche == -77, NA, age_menarche))
```



###CAIDE

```{r}
caide<-read_excel("G:/My Drive/Data/ALFA+/CVR/20220325_APlus_CAIDE.xlsx")

caide<-caide%>%
  filter(Visit == 1)%>%
  dplyr::select(IdParticipante, CAIDE_scr_20M1)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))

master_nd<-left_join(master_nd, caide)
```

### men's reproductive history
```{r}

men_repro<-read_excel("G:/My Drive/Data/ALFA+/BBDD_chldren_prealfa_clara.xlsx")%>%
  dplyr::select(ID, total)%>%
  rename(subject = ID, 
         kids_total = total)%>%
  mutate(subject = factor(subject))
  
master_nd<-left_join(master_nd, men_repro)

#replacing unequal values between our women only variable and the variable just added by the values in our women only variable, which is definitely errorless
#then replacing NA with 0, since the variable just included is longform with entries only for participants with kids

master_nd<-master_nd%>%
  mutate(kids_total = ifelse(kids_total != preg_termino & !is.na(preg_termino), preg_termino, kids_total),
         kids_total = replace_na(kids_total, 0))%>%
  dplyr::select(-preg_termino)

```


###BP + physical activity
F13_011C	Índice de masa corporal (IMC)
BP_SIST_2	TA sistólica (mmHg) - TOMA-2
BP_DIAST_2	TA diastólica (mmHg) - TOMA-2
AL_09_038C	Minnesota Resultado Actividad Física según tipo de ejercico. 0 is inactive, 1 is active

```{r}
risks<-read_excel("G:/My Drive/Data/ALFA+/20210517_APlus_General_Clinical.xlsx")%>%
  dplyr::select(IdParticipante, BP_SIST_2, BP_DIAST_2, AL_09_038C, F13_011C)%>%
  mutate(F13_011C = as.numeric(F13_011C),
         BP_SIST_2 = as.numeric(BP_SIST_2),
         BP_DIAST_2 = as.numeric(BP_DIAST_2))%>%
  rename(subject=IdParticipante,
         bmi = F13_011C,
         coef_physact = AL_09_038C)%>%
  mutate(subject = factor(subject))

master_nd<-left_join(master_nd, risks)

#two participanta only one measure, inputing manually
master_nd$BP_SIST_2[master_nd$subject == 66240] <- 137
master_nd$BP_DIAST_2[master_nd$subject == 66240] <- 70
master_nd$BP_SIST_2[master_nd$subject ==10601] <-133
master_nd$BP_DIAST_2[master_nd$subject ==10601] <-68
```

###cholesterol
```{r}
cholesterol <- read_delim("G:/My Drive/Data/ALFA+/biomarkers/Laboratorio_v1.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

cholesterol<-cholesterol%>%
  dplyr::select(IdParticipante, cholesterol)%>%
  mutate(IdParticipante=factor(IdParticipante))%>%
  rename(subject = IdParticipante)

master_nd<-left_join(master_nd, cholesterol)
```

###smoking status
```{r}
smoke <- read_delim("G:/My Drive/Data/ALFA+/CVR/smoking status alfa+ v1.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(Fuma2 = col_factor(levels = c("0", 
        "1", "2")), ...3 = col_skip(), ...4 = col_skip(), 
        ...5 = col_skip()), trim_ws = TRUE)

# 0= never smoke
# 1= current smoker or ex-smoker since < 1 year
# 2= exsmoker since >5 years

smoke<-smoke%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))

master_nd<-left_join(master_nd, smoke)

master_nd$Fuma2[master_nd$subject == "11849"]<-0
```

###medical antecedents
```{r}
#antecedentes patogenicos
AP<-read_excel("G:/My Drive/Data/ALFA+/CVR/20210517_APlus-AP_copy.xlsx")

AP<-AP%>%
  filter(!is.na(IdParticipante))%>%
  dplyr::select(IdParticipante, AP_CVhbp, AP_METDM, AP_METdislip)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject=factor(subject),
         AP_CVhbp = replace_na(AP_CVhbp, 0),
         AP_METDM = replace_na(AP_METDM, 0),
         AP_METdislip = replace_na(AP_METdislip, 0))

master_nd<-left_join(master_nd, AP)
```

###Medication
```{r}
#medications
meds<-read_excel("G:/My Drive/Data/ALFA+/CVR/ALFA+_v1_meds.xls")
meds<-meds%>%
  dplyr::select(idparticipante, anti_hta, `anti_dyslip (cholesterol)`, anti_diab)%>%
  rename(subject = idparticipante,
         anti_dyslip = `anti_dyslip (cholesterol)`)%>%
  mutate(subject = factor(subject),
         anti_hta = replace_na(anti_hta, 0),
         anti_dyslip = replace_na(anti_dyslip, 0),
         anti_diab = replace_na(anti_diab, 0),)


master_nd<-left_join(master_nd, meds)

master_nd<-master_nd%>%
  mutate(anti_hta = replace_na(anti_hta, 0),
         anti_dyslip = replace_na(anti_dyslip, 0),
         anti_diab = replace_na(anti_diab, 0))

```


## social factors

11116	F01_050B	Nivel socioeconómico referido por el participante:
1	Nivel económico bajo
2	Nivel económico medio
3	Nivel económico alto
-77	No Sabe
-88	No Evaluado

11119	DEMOG_famincomecat	Ingresos brutos anuales del hogar (unidad familiar) participante.
0	No tiene ingresos
1	Menos de 6.000€
2	De 6.001 a 9.000€
3	De 9.001 a 12.000€
4	De 12.001 a 18.000€
5	De 18.001 a 30.000€
6	De 30.001 a 54.000€
-66	No quiere responder
7	De 54.001 a 90.000€
-77	No lo sabe
8	Más de 90.000€
-88	No Evaluado

11083	F01_015B	Estado civil: 
1	Soltero/a
2	Casado/a o Pareja de hecho/estable
3	Separado/a o Divorciado/a
4	Viudo/a
5	Otro estado civil
-88	No Evaluado

,
         DEMOG_famincomecat = case_when(DEMOG_famincomecat == -77 ~ NA,
                                        DEMOG_famincomecat == -66 ~ NA,
                                        DEMOG_famincomecat == -88 ~ NA,
                                        T ~ DEMOG_famincomecat)
```{r}
socio<-read_delim("G:/My Drive/Data/ALFA+/socio_var/Alfa Plus Long 2024_1020_DEMOG_20241219102604(in).csv", delim = ";", escape_double = FALSE, trim_ws = TRUE, skip=1)%>%
  filter(Visita == "V2"| Visita == "V3")%>%
  dplyr::select(IdParticipante, Visita, DEMOG_famincomecat, F01_015B)%>%
  rename(estado_civil = F01_015B)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject),
         DEMOG_famincomecat = case_when(DEMOG_famincomecat == -77 ~ NA,
                                        DEMOG_famincomecat == -66 ~ NA,
                                        DEMOG_famincomecat == -88 ~ NA,
                                        T ~ DEMOG_famincomecat))%>%
  group_by(subject)%>%
  mutate(ingreso_hogar = if(sum(is.na(DEMOG_famincomecat)) == 1 & n() == 2) {
    # If exactly one NA and two observations per participant
    # Replace all values with the non-NA value
    rep(DEMOG_famincomecat[!is.na(DEMOG_famincomecat)][1], n())
  } else {
    # Otherwise keep values as they are
    DEMOG_famincomecat
  })%>%
  ungroup()%>%
  filter(Visita == "V2")%>%
  dplyr::select(-Visita, -DEMOG_famincomecat)



master_nd<-left_join(master_nd, socio)
```

## Occupational attainment
0	No cualificada (incluye sus labores)
1	Cualificado manual
2	Cualificado no manual (incluye secretariado técnico)
3	Profesional (estudios superiores)
4	Directivo

group 0 and 1 together; rationale is that both have no 

```{r}
occupation<- read_excel("G:/My Drive/Data/ALFA+/socio_var/occupational attainment.xlsx", skip = 1)%>%
  filter(Visita == "V1")%>%
  dplyr::select(IdParticipante, F02_4_012B)%>%
  rename(subject = IdParticipante, 
         occupation = F02_4_012B)%>%
  mutate(subject = factor(subject),
         occupation = factor(occupation))


master_nd<-left_join(master_nd, occupation)

```

## SNAC
value reflects the total amount of adverse/stressful events in one's life
```{r}
snac<-read_excel("G:/My Drive/Data/ALFA+/20210517_APlus_General_Clinical.xlsx")%>%
  dplyr::select(IdParticipante, AO_03_044C)%>%
  rename(subject = IdParticipante,
         SNAC = AO_03_044C)%>%
  mutate(subject = factor(subject),
         SNAC = factor(SNAC))

master_nd<-left_join(master_nd, snac)
```

##APOE status

```{r}
apoe<-read_delim("G:/My Drive/Data/ALFA+/biomarkers/20201007_APlus_Biomarcadores - corregido.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)%>%
  dplyr::select(IdParticipante, APOE_binary)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject),
         APOE_binary = factor(APOE_binary))%>%
  unique()

master_nd<-left_join(master_nd, apoe)
```

## cognitive scores - not included atm

This dataset only contains V1 and V2 scores for now.

1.	FCSRT Total Immediate recall (0-48), Z_FCSRTI_RT -> word recall based on semantic cues
2.	WMS logical memory Total Delayed recall (0-50) Z_aWMS_DR_RS -> narrative memory
3.	WAIS- Coding (0-135) Z_Wclave_total -> attention/processing
4.	Semantic Fluency (animals in 1 minute), Z_FS_totalcnswr -> semantic memory


```{r}
# cog_v1<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 1 - Cognition Z Scores.xlsx")%>%
#   dplyr::select(-Gender, -Years_Edu, -Age_V1S1)%>%
#   rename(subject = IdParticipante,
#          Z_PACC_v1 = Z_PACC_Composite_1.2._v1)
# 
# 
# cog_v2<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 2 - Cognition Z Scores.xlsx")%>%
#   dplyr::select(-Gender, -Years_Edu, -Age_V1S1, -V1S1_dt)%>%
#   rename(subject = IdParticipante,
#          Z_PACC_v2 = Z_PACC_Composite_1.2._V2)
# 
# cog_all<-full_join(cog_v1, cog_v2)
# 
# #create time difference variable for adjusting in models, and create the score difference between V1 and V2
# 
# cog_all<-cog_all%>%
#   mutate(V1S1_dt = as.Date(V1S1_dt, format = "%d/%m/%Y"),
#          V2S1_dt = as.Date(V2S1_dt, format = "%d/%m/%Y"),
#          diff_time = difftime(V2S1_dt, V1S1_dt, units = c("days")),
#          diff_PACC = Z_PACC_v2 - Z_PACC_v1,
#          diff_Attention = Z_Attention_Composite_V2 - Z_Attention_Composite_v1,
#          diff_Executive = Z_Executive_Composite_V2 - Z_Executive_Composite_v1,
#          diff_Language = Z_Language_Composite_V2 - Z_Language_Composite_v1,
#          diff_Visual = Z_Visual_Composite_V2 - Z_Visual_Composite_v1,
#          diff_Memory = Z_Memory_Composite_V2 - Z_Memory_Composite_v1)
# 
# 
# 
# 
# master_nd<-left_join(master_nd, cog_all)
# master_nd$diff_time<-as.numeric(master_nd$diff_time)

```




#handling missing data
##check
```{r}
master_nd%>%
  md.pattern(., rotate.names = T)

aggr_plot <- VIM::aggr(master_nd[master_nd$sex== 2,], numbers = T, sortVars= T)
#this tells us that all AP variables, plus age at first birth have missing rates that are too high. Discuss this tomorrow


```

##fix 

Replacing NAs: predicitve means matching is used with 5 imputations for each value. I deem this necessary and useful to keep the little power we have in the linear regression since losing data because of covariates suck. NAs are missing at random

a new dataset is created with the new variables that have been imputed. New variables wear the suffix "_pmm" and old ones were delected from the new dataset

threshold for imputing is 10% missing data, fairly conservative

outcome variables aren't being computed, only covariates



```{r}


#cholesterol

mice_dat<-master_nd%>%
  dplyr::select(cholesterol, subject, age_v1, sex, BP_SIST_2, bmi, education , APOE_binary)%>%
  distinct()

ini <- mice(mice_dat,maxit=0)
pred1 <- ini$predictorMatrix
pred1[,'subject'] <- 0
imp_multi <- mice(mice_dat, pred = pred1, m = 5, method = "pmm")  # Impute missing values multiple times
data_imp_multi_all <- complete(imp_multi,       # Store multiply imputed data
                           "repeated",
                           include = TRUE)

cholesterol_pmm<-data_imp_multi_all%>%
  dplyr::select(2:7)%>%
  mutate(cholesterol_pmm = rowMeans(across(starts_with("cholesterol."))))%>%
  rename(subject=subject.0)%>%
  mutate(subject = factor(subject))

master_nd_pmm<-master_nd
master_nd_pmm<-left_join(master_nd_pmm, cholesterol_pmm[,c(6, 7)])

## age_menopause

mice_dat<-master_nd%>%
  filter(sex == 2)%>%
  dplyr::select(age_menopause, subject, age_v1, sex, BP_SIST_2, bmi, education , APOE_binary)%>%
  distinct()

ini <- mice(mice_dat,maxit=0)
pred1 <- ini$predictorMatrix
pred1[,'subject'] <- 0
imp_multi <- mice(mice_dat, pred = pred1, m = 5, method = "pmm")  # Impute missing values multiple times
data_imp_multi_all <- complete(imp_multi,       # Store multiply imputed data
                           "repeated",
                           include = TRUE)

age_menopause_pmm<-data_imp_multi_all%>%
  dplyr::select(2:7)%>%
  mutate(age_menopause_pmm = rowMeans(across(starts_with("age_menopause."))))%>%
  rename(subject=subject.0)%>%
  mutate(subject = factor(subject))


master_nd_pmm<-left_join(master_nd_pmm, age_menopause_pmm[,c(6, 7)])


master_nd_pmm%>%
  md.pattern(., rotate.names = T)

aggr_plot <- VIM::aggr(master_nd_pmm, numbers = T, sortVars= T)

```



#create vars for analysis

## reproductive var
```{r}
#recompute reproductive_span variable
master_nd_pmm<-master_nd_pmm%>%
  mutate(repro_span = age_menopause_pmm - age_menarche,
         kids_dicho = factor(ifelse(kids_total == 0, 0, 1)),
         parous_group = case_when(
  kids_total >= 2 ~ "2+",
  kids_total == 1 ~ "1",
  TRUE ~ "0"))


```


##CVR variables
```{r}
master_nd_pmm<-master_nd_pmm%>%
  mutate(hypertension = ifelse(BP_SIST_2>=140|BP_DIAST_2>=90| AP_CVhbp == 1 |anti_hta==1, 1, 0),
         high_chol = ifelse(cholesterol_pmm>=200 |AP_METdislip == 1 | anti_dyslip ==1, 1, 0),
         diabetes = ifelse(AP_METDM == 1 |anti_diab ==1, 1, 0))

master_nd_pmm<-master_nd_pmm%>%
  mutate(hypertension = factor(hypertension),
         high_chol = factor(high_chol),
         diabetes = factor(diabetes))%>%
  dplyr::select(-AP_METDM, -AP_METdislip, -AP_CVhbp, -anti_dyslip, -anti_hta, -anti_diab)
```


#Social variables
to adjust for the very unequal amount of observation per bin
```{r}
master_nd_pmm<-master_nd_pmm%>%
  mutate(new_ingreso = case_when(ingreso_hogar <=5 ~ "low",
                                 ingreso_hogar == 6 ~ "medium",
                                 ingreso_hogar>6 ~ "high"),
         new_ingreso2 = case_when(ingreso_hogar <5 ~ "low",
                                 ingreso_hogar == 5 ~ "medium",
                                 ingreso_hogar == 6 ~ "high",
                                 ingreso_hogar>6 ~ "very high"))

master_nd_pmm<-master_nd_pmm%>%
  mutate(new_ingreso2 = factor(new_ingreso2))


```
# Save output
```{r}
  

aggr_plot <- VIM::aggr(master_nd_pmm, numbers = T, sortVars= T)

write.csv(master_nd_pmm, file = "G:/My Drive/Data/ALFA+/ALFA_dataset_basic.csv", row.names = F)
```
























