---
title: "Basic ALFA+ dataset"
author: "Clazz"
date: "2025-05-27"
output: html_document
---


```{r}
library(readxl)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("psych")
library(psych)
#install.packages("tidyverse")
library(tidyverse)
#install.packages("ggpubr")
library(ggpubr)
#install.packages("broom")
library(broom)
#install.packages("ggeffects")
library(ggeffects)
#install.packages("MASS")
library(MASS)
#install.packages("mice")
library(mice)
#install.packages("VIM")
library(VIM)
#install.packages("car")
library(car)
#install.packages("sjPlot")
library(sjPlot)
#install.packages("caret")
library(caret)
#install.packages("bestNormalize")
library(bestNormalize)
#install.packages("ggrain")
library(ggrain)
#install.packages("emmeans")
library(emmeans)
#install.packages("skimr")
library(skimr)
library(glmnet)
#install.packages("bullseye")
library(bullseye)
#install.packages("lavaan")
library(lavaan)
#install.packages("brms")
library(brms)
#install.packages("mgcv")
library(mgcv)
library(splines)
#install.packages("fitdistrplus")
library(fitdistrplus)
#install.packages("patchwork")
library(patchwork)
```

This code processes ALFA+ cohort data into a main dataset. variable specs can be found in the READ_ME file


#making the dataset

##alfa visits basic dataset

```{r}

master_nd<- read_excel("G:/My Drive/Data/ALFA+/data_raw/list_alfa_subjects_v1.xlsx")%>%
  dplyr::select(ID, sex, age_v1,education)%>%
  rename(subject = ID)%>%
  mutate(subject = factor(subject),
         sex= factor(sex),
         age_v1 = as.numeric(age_v1))
bd<-read_excel("G:/My Drive/Data/ALFA+/data_raw/20210517_APlus_General_Clinical.xlsx")%>%
  rename(subject = IdParticipante)%>%
   mutate(subject = factor(subject))%>%
  dplyr::select(subject, DOB)%>%
  unique()

master_nd<-left_join(master_nd, bd)

```



##TIV
```{r}
TIV<- read_excel("G:/My Drive/Data/ALFA+/data_raw/neuroimaging/TIV_ALFA+_V1.xlsx")

TIV<-TIV%>%
  dplyr::select(subject, value)%>%
  mutate(subject = factor(subject))%>%
  rename(TIV = value)

#add TIV
master_nd<-left_join(master_nd, TIV)

```

##read in the modifiable risk factors 

###women's reproductive history
I want to find out which of the women had more than 4 kids that they changed...
```{r}

#removing women who reported pregnancy related diseases 

women_preg<-read_delim("G:/My Drive/Data/ALFA+/data_raw/reproductive_var/Alfa Plus Long 2025_1054_AF_CRF_AF_embarazos_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  filter(Visita == "V2")%>%
  dplyr::select(IdParticipante, AF_deliverypat, AF_deliveryage)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))%>%
  group_by(subject)%>%
  mutate(AF_deliverypat = max(AF_deliverypat),
         AF_deliveryage = min(AF_deliveryage))%>%
  ungroup()%>%
  unique()%>%
  rename(Age_at_first_birth = AF_deliveryage)

women_preg%>%
  filter(AF_deliverypat !=0)

master_nd<-left_join(master_nd, women_preg)

master_nd<-master_nd%>%
  mutate(AF_deliverypat = replace_na(AF_deliverypat, 0))%>%
  filter(AF_deliverypat == 0)%>%
  dplyr::select(-AF_deliverypat)



#age menopause + abortions + menarche 

menopause<-read_delim("G:/My Drive/Data/ALFA+/data_raw/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, F06_2_017, F06_2_019, F06_2_018, F06_2_015)%>%
  rename(subject = IdParticipante,
         age_menopause = F06_2_017, 
         abortions = F06_2_019,
         preg_termino = F06_2_018,
         age_menarche = F06_2_015)%>%
  filter(!if_all(age_menopause:age_menarche, is.na))%>%
  group_by(subject)%>%
  mutate(age_menopause = max(age_menopause),
         abortions = max(abortions),
         age_menarche = min(age_menarche))%>%
  ungroup()%>%
  unique()%>%
  mutate(subject = factor(subject),
         age_menopause = ifelse(age_menopause == 0, NA, age_menopause))

master_nd<-left_join(master_nd, menopause)

cause_menopause<-read_delim("G:/My Drive/Data/ALFA+/data_raw/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, AF_menopcs)%>%
  rename(cause_menopause = AF_menopcs)%>%
  mutate(cause_menopause = ifelse(cause_menopause == -77 | cause_menopause == -66, NA, cause_menopause))%>%
  drop_na(cause_menopause)%>%
  mutate(cause_menopause = factor(ifelse(cause_menopause == 3, "non-natural", "natural")))%>%
  unique()%>%
  filter(!(IdParticipante == "11397" & cause_menopause == "natural"))%>%
  filter(!(IdParticipante == "55036" & cause_menopause == "natural"))%>%
  filter(!(IdParticipante == "11882" & cause_menopause == "natural"))#found duplicate entries becuse participants answers change

cause_early_menopause<-read_delim("G:/My Drive/Data/ALFA+/data_raw/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, AF_earlymenopcs, AF_horm_sust)%>%
  mutate(AF_earlymenopcs = factor(AF_earlymenopcs))
  
cause_early_menopause%>%
  filter(AF_earlymenopcs == 2)%>%
  unique()%>%
  pull(IdParticipante)

cause_menopause$cause_menopause[cause_menopause$IdParticipante == 10225]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 10388]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 11111]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 11114]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 11557]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 44193]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 55036]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 55274]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 66301]<-"non-natural"
cause_menopause$cause_menopause[cause_menopause$IdParticipante == 77191]<-"non-natural"

cause_menopause<-cause_menopause%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))

menopause%>%
  drop_na(age_menopause)

master_nd<-left_join(master_nd, cause_menopause)

# HRT
hrt<-read_delim("G:/My Drive/Data/ALFA+/data_raw/reproductive_var/Alfa Plus Long 2025_1054_AF_20250304095809(in).csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE, skip = 1)%>%
  dplyr::select(IdParticipante, AF_horm_sust, Visita)%>%
  mutate(AF_horm_sust = factor(AF_horm_sust),
         IdParticipante = factor(IdParticipante))%>%
  rename(HRT = AF_horm_sust,
         subject = IdParticipante)%>%
  dplyr::select(-Visita)%>%
  unique()%>%
  na.omit(HRT)%>%
  group_by(subject)%>%
  filter(!(n() > 1 & HRT == 0)) %>%
  ungroup()

master_nd<-left_join(master_nd, hrt)

#replacing the no answers values with NA
master_nd<-master_nd%>%
  mutate(age_menopause = ifelse(age_menopause == -77, NA, age_menopause),
         age_menarche = ifelse(age_menarche == -77, NA, age_menarche))

```



###CAIDE

```{r}
caide<-read_excel("G:/My Drive/Data/ALFA+/data_raw/CVR/20220325_APlus_CAIDE.xlsx")

caide<-caide%>%
  filter(Visit == 1)%>%
  dplyr::select(IdParticipante, CAIDE_scr_20M1)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))

master_nd<-left_join(master_nd, caide)
```

### men's reproductive history
```{r}

men_repro<-read_excel("G:/My Drive/Data/ALFA+/data_raw/reproductive_var/BBDD_chldren_prealfa_clara.xlsx")%>%
  dplyr::select(ID, total)%>%
  rename(subject = ID, 
         kids_total = total)%>%
  mutate(subject = factor(subject))
  
master_nd<-left_join(master_nd, men_repro)

#replacing unequal values between our women only variable and the variable just added by the values in our women only variable, which is definitely errorless
#then replacing NA with 0, since the variable just included is longform with entries only for participants with kids

master_nd<-master_nd%>%
  mutate(kids_total = ifelse(kids_total != preg_termino & !is.na(preg_termino), preg_termino, kids_total),
         kids_total = replace_na(kids_total, 0))%>%
  dplyr::select(-preg_termino)

```


###BP + physical activity
F13_011C	Índice de masa corporal (IMC)
BP_SIST_2	TA sistólica (mmHg) - TOMA-2
BP_DIAST_2	TA diastólica (mmHg) - TOMA-2
AL_09_038C	Minnesota Resultado Actividad Física según tipo de ejercico. 0 is inactive, 1 is active

```{r}
risks<-read_excel("G:/My Drive/Data/ALFA+/data_raw/20210517_APlus_General_Clinical.xlsx")%>%
  dplyr::select(IdParticipante, BP_SIST_2, BP_DIAST_2, AL_09_038C, F13_011C)%>%
  mutate(F13_011C = as.numeric(F13_011C),
         BP_SIST_2 = as.numeric(BP_SIST_2),
         BP_DIAST_2 = as.numeric(BP_DIAST_2))%>%
  rename(subject=IdParticipante,
         bmi = F13_011C,
         coef_physact = AL_09_038C)%>%
  mutate(subject = factor(subject))

master_nd<-left_join(master_nd, risks)

#two participanta only one measure, inputing manually
master_nd$BP_SIST_2[master_nd$subject == 66240] <- 137
master_nd$BP_DIAST_2[master_nd$subject == 66240] <- 70
master_nd$BP_SIST_2[master_nd$subject ==10601] <-133
master_nd$BP_DIAST_2[master_nd$subject ==10601] <-68
```


###smoking status
```{r}
smoke <- read_delim("G:/My Drive/Data/ALFA+/data_raw/CVR/smoking status alfa+ v1.csv", 
    delim = ";", escape_double = FALSE, col_types = cols(Fuma2 = col_factor(levels = c("0", 
        "1", "2")), ...3 = col_skip(), ...4 = col_skip(), 
        ...5 = col_skip()), trim_ws = TRUE)

# 0= never smoke
# 1= current smoker or ex-smoker since < 1 year
# 2= exsmoker since >5 years

smoke<-smoke%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject))

master_nd<-left_join(master_nd, smoke)

master_nd$Fuma2[master_nd$subject == "11849"]<-0
```






## social factors

11116	F01_050B	Nivel socioeconómico referido por el participante:
1	Nivel económico bajo
2	Nivel económico medio
3	Nivel económico alto
-77	No Sabe
-88	No Evaluado

11119	DEMOG_famincomecat	Ingresos brutos anuales del hogar (unidad familiar) participante.
0	No tiene ingresos
1	Menos de 6.000€
2	De 6.001 a 9.000€
3	De 9.001 a 12.000€
4	De 12.001 a 18.000€
5	De 18.001 a 30.000€
6	De 30.001 a 54.000€
-66	No quiere responder
7	De 54.001 a 90.000€
-77	No lo sabe
8	Más de 90.000€
-88	No Evaluado

11083	F01_015B	Estado civil: 
1	Soltero/a
2	Casado/a o Pareja de hecho/estable
3	Separado/a o Divorciado/a
4	Viudo/a
5	Otro estado civil
-88	No Evaluado

,
         DEMOG_famincomecat = case_when(DEMOG_famincomecat == -77 ~ NA,
                                        DEMOG_famincomecat == -66 ~ NA,
                                        DEMOG_famincomecat == -88 ~ NA,
                                        T ~ DEMOG_famincomecat)
```{r}
socio<-read_delim("G:/My Drive/Data/ALFA+/data_raw/socio_var/Alfa Plus Long 2024_1020_DEMOG_20241219102604(in).csv", delim = ";", escape_double = FALSE, trim_ws = TRUE, skip=1)%>%
  filter(Visita == "V2"| Visita == "V3")%>%
  dplyr::select(IdParticipante, Visita, DEMOG_famincomecat, F01_015B)%>%
  rename(estado_civil = F01_015B)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject),
         DEMOG_famincomecat = case_when(DEMOG_famincomecat == -77 ~ NA,
                                        DEMOG_famincomecat == -66 ~ NA,
                                        DEMOG_famincomecat == -88 ~ NA,
                                        T ~ DEMOG_famincomecat))%>%
  group_by(subject)%>%
  mutate(ingreso_hogar = if(sum(is.na(DEMOG_famincomecat)) == 1 & n() == 2) {
    # If exactly one NA and two observations per participant
    # Replace all values with the non-NA value
    rep(DEMOG_famincomecat[!is.na(DEMOG_famincomecat)][1], n())
  } else {
    # Otherwise keep values as they are
    DEMOG_famincomecat
  })%>%
  ungroup()%>%
  filter(Visita == "V2")%>%
  dplyr::select(-Visita, -DEMOG_famincomecat)



master_nd<-left_join(master_nd, socio)
```

## Occupational attainment
0	No cualificada (incluye sus labores)
1	Cualificado manual
2	Cualificado no manual (incluye secretariado técnico)
3	Profesional (estudios superiores)
4	Directivo

group 0 and 1 together; rationale is that both have no 

```{r}
occupation<- read_excel("G:/My Drive/Data/ALFA+/data_raw/socio_var/occupational attainment.xlsx", skip = 1)%>%
  filter(Visita == "V1")%>%
  dplyr::select(IdParticipante, F02_4_012B)%>%
  rename(subject = IdParticipante, 
         occupation = F02_4_012B)%>%
  mutate(subject = factor(subject),
         occupation = factor(occupation))


master_nd<-left_join(master_nd, occupation)

```

## SNAC
value reflects the total amount of adverse/stressful events in one's life
```{r}
# snac<-read_excel("G:/My Drive/Data/ALFA+/data_raw/20210517_APlus_General_Clinical.xlsx")%>%
#   dplyr::select(IdParticipante, AO_03_044C)%>%
#   rename(subject = IdParticipante,
#          SNAC = AO_03_044C)%>%
#   mutate(subject = factor(subject),
#          SNAC = factor(SNAC))
# 
# master_nd<-left_join(master_nd, snac)
```

##APOE status

```{r}
apoe<-read_delim("G:/My Drive/Data/ALFA+/data_raw/fluid_biomarkers/20201007_APlus_Biomarcadores - corregido.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)%>%
  dplyr::select(IdParticipante, APOE_binary)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = factor(subject),
         APOE_binary = factor(APOE_binary))%>%
  unique()

master_nd<-left_join(master_nd, apoe)
```

#handling missing data

Replacing NAs: Not doing this method anymore, apparently flawed. data needs to be imputed at least 10 times, for the specific regression used. 
Example:

-> Get all imputed datasets in long format
data_imp_long <- complete(imp_multi, "long", include = TRUE)

-> Run analysis within each imputation
library(broom)
results <- data_imp_long %>%
  filter(.imp > 0) %>%  # Exclude original data
  group_by(.imp) %>%
  do(tidy(lm(outcome ~ cholesterol + age_v1, data = .))) %>%
  ungroup()
  
fit <- with(imp_multi, lm(outcome ~ covariate1 + covariate2 + covariate3 + covariate4 + covariate5))
pooled <- pool(fit)
summary(pooled)


#create vars for analysis

## reproductive var
```{r}
#recompute reproductive_span variable
master_nd<-master_nd%>%
  mutate(repro_span = age_menopause - age_menarche,
         kids_dicho = factor(ifelse(kids_total == 0, 0, 1)),
         parous_group = case_when(
  kids_total >= 2 ~ "2+",
  kids_total == 1 ~ "1",
  TRUE ~ "0"))

```


#Other (not processed currently)
cholesterol procssing is missing too
```{r}
# master_nd<-master_nd%>%
#   mutate(hypertension = ifelse(BP_SIST_2>=140|BP_DIAST_2>=90| AP_CVhbp == 1 |anti_hta==1, 1, 0),
#          high_chol = ifelse(cholesterol_pmm>=200 |AP_METdislip == 1 | anti_dyslip ==1, 1, 0),
#          diabetes = ifelse(AP_METDM == 1 |anti_diab ==1, 1, 0))
# 
# master_nd<-master_nd%>%
#   mutate(hypertension = factor(hypertension),
#          high_chol = factor(high_chol),
#          diabetes = factor(diabetes))%>%
#   dplyr::select(-AP_METDM, -AP_METdislip, -AP_CVhbp, -anti_dyslip, -anti_hta, -anti_diab)
```

###medical antecedents
```{r}
#antecedentes patogenicos
# AP<-read_excel("G:/My Drive/Data/ALFA+/data_raw/CVR/20210517_APlus-AP_copy.xlsx")
# 
# AP<-AP%>%
#   filter(!is.na(IdParticipante))%>%
#   dplyr::select(IdParticipante, AP_CVhbp, AP_METDM, AP_METdislip)%>%
#   rename(subject = IdParticipante)%>%
#   mutate(subject=factor(subject),
#          AP_CVhbp = replace_na(AP_CVhbp, 0),
#          AP_METDM = replace_na(AP_METDM, 0),
#          AP_METdislip = replace_na(AP_METdislip, 0))
# 
# master_nd<-left_join(master_nd, AP)
```

###Medication
```{r}
#medications
# meds<-read_excel("G:/My Drive/Data/ALFA+/data_raw/CVR/ALFA+_v1_meds.xls")
# meds<-meds%>%
#   dplyr::select(idparticipante, anti_hta, `anti_dyslip (cholesterol)`, anti_diab)%>%
#   rename(subject = idparticipante,
#          anti_dyslip = `anti_dyslip (cholesterol)`)%>%
#   mutate(subject = factor(subject),
#          anti_hta = replace_na(anti_hta, 0),
#          anti_dyslip = replace_na(anti_dyslip, 0),
#          anti_diab = replace_na(anti_diab, 0),)
# 
# 
# master_nd<-left_join(master_nd, meds)
# 
# master_nd<-master_nd%>%
#   mutate(anti_hta = replace_na(anti_hta, 0),
#          anti_dyslip = replace_na(anti_dyslip, 0),
#          anti_diab = replace_na(anti_diab, 0))

```

# Save output
```{r}
  

aggr_plot <- VIM::aggr(master_nd, numbers = T, sortVars= T)

write.csv(master_nd, file = "G:/My Drive/Data/ALFA+/data_cleaned/ALFA_dataset_basic.csv", row.names = F)
```
























