---
title: "Poster AAIC 2025"
author: "Clazz"
date: "2025-07-01"
output: html_document
---

```{r}
library(readxl)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("psych")
library(psych)
#install.packages("tidyverse")
library(tidyverse)
#install.packages("ggpubr")
library(ggpubr)
#install.packages("broom")
library(broom)
#install.packages("ggeffects")
library(ggeffects)
#install.packages("MASS")
library(MASS)
#install.packages("PResiduals")
library(PResiduals)
#install.packages("mice")
library(mice)
#install.packages("VIM")
library(VIM)
#install.packages("car")
library(car)
#install.packages("sjPlot")
library(sjPlot)
#install.packages("caret")
library(caret)
#install.packages("bestNormalize")
library(bestNormalize)
#install.packages("ggrain")
library(ggrain)
#install.packages("emmeans")
library(emmeans)
#install.packages("skimr")
library(skimr)
library(glmnet)
#install.packages("bullseye")
library(bullseye)
#install.packages("lavaan")
library(lavaan)
#install.packages("brms")
library(brms)
#install.packages("mgcv")
library(mgcv)
library(splines)
#install.packages("fitdistrplus")
library(fitdistrplus)
#install.packages("patchwork")
library(patchwork)
library(trafo)
```

#Read in data
```{r}
wmh<-read_delim("G:/My Drive/Data/ALFA+/WMH/wmh_processed_regional_lobes.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)%>%
  filter(Visit == "V1")%>%
  dplyr::select(-Visit)

cognition<-read_excel("G:/My Drive/Data/ALFA+/Cognition/PACCv1_to_v3(incomplete).xlsx")%>%
  dplyr::select(IdParticipante, Visit, Z_PACC, FechaResultado)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = as.numeric(subject))%>%
  filter(Visit != 3)%>%
  pivot_wider(names_from = Visit, values_from = c(Z_PACC, FechaResultado))%>%
  mutate(PACC_diff = Z_PACC_2 - Z_PACC_1,
         time_diff = as.numeric(difftime(FechaResultado_2, FechaResultado_1, units= c("days"))))%>%
  dplyr::select(subject, PACC_diff, time_diff)

composites1<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 1 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -Z_PACC_Composite_1.2._v1 )%>%
  rename(subject = IdParticipante)

composites2<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 2 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -V2S1_dt, -Z_PACC_Composite_1.2._V2 )%>%
  rename(subject = IdParticipante)

composite_all<-full_join(composites1, composites2)%>%
  mutate(diff_Attention = Z_Attention_Composite_V2 - Z_Attention_Composite_v1,
         diff_Memory = Z_Memory_Composite_V2 - Z_Memory_Composite_v1,
         diff_Executive = Z_Executive_Composite_V2 - Z_Executive_Composite_v1,
         diff_Language = Z_Language_Composite_V2 - Z_Language_Composite_v1,
         diff_Visual = Z_Visual_Composite_V2 - Z_Visual_Composite_v1,
         subject = as.numeric(subject))%>%
  dplyr:: select(subject, diff_Attention, diff_Memory, diff_Executive, diff_Language, diff_Visual)


biomarkers<-read_delim("G:/My Drive/Data/ALFA+/biomarkers/biomarkers_CSF_processed.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)%>%
  filter(Visit == "V1")%>%
  dplyr::select(subject, A, AT)

data_wmh<-read_delim("G:/My Drive/Data/ALFA+/ALFA_dataset_basic.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)%>%
  left_join(., wmh)%>%
  left_join(., cognition)%>%
  left_join(., biomarkers)%>%
  left_join(., composite_all)%>%
  mutate(APOE_binary = factor(APOE_binary),
         A = factor(A))

data_wmh_w<-data_wmh%>%
  filter(sex == 2)
```

#sanity checks

```{r}
data_wmh%>%
  filter(bilateral_region == "occipital")%>%
  pull(volume_wmh)%>%
  hist()


```
#Model 0 - wmh ~kids

```{r}
mod_1<-lm(volume_wmh ~ kids_total  + APOE_binary + education+ TIV+ age_v1 , data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_2<-lm(volume_wmh ~ kids_total + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "parietal",])
mod_3<-lm(volume_wmh ~ kids_total  + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])
mod_4<-lm(volume_wmh ~ kids_total  + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "occipital",])
mod_5<-lm(volume_wmh ~ kids_total  + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])

summary(trafo_lm(mod_1))
summary(trafo_lm(mod_2))
summary(trafo_lm(mod_3))
summary(trafo_lm(mod_4))
summary(trafo_lm(mod_5))

p_values <- c(0.043811 ,0.233419  ,  0.0265 ,0.674804  ,  0.0155 )
#None survive, basal ganglia becomes trend

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

data_wmh_w%>%
  filter(bilateral_region %in% c("temporal", "frontal", "basal ganglia"))%>%
  ggplot(aes(x= kids_total, y= volume_wmh))+
  geom_jitter()+
  geom_smooth(method = "lm")+
  facet_wrap(~bilateral_region, scale = "free")
  
```


#Model 1 - PACC ~ WMH

```{r}
mod_1<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education  + TIV+ age_v1 +A, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_2<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1+A, data = data_wmh_w[data_wmh_w$bilateral_region == "parietal",])
mod_3<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1+A, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])
mod_4<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1+A, data = data_wmh_w[data_wmh_w$bilateral_region == "occipital",])
mod_5<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1+A, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])

summary(trafo_lm(mod_1))
summary(trafo_lm(mod_2))
summary(trafo_lm(mod_3))
summary(trafo_lm(mod_4))
summary(trafo_lm(mod_5))


```

no effect of WMH on PACC change between v1 and v2. No trend of A status

# Model 2: PACC ~ childbirth 

```{r}
mod_1<-lm(PACC_diff ~ kids_total + time_diff + APOE_binary + education + TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_2<-lm(PACC_diff ~ kids_dicho + time_diff + APOE_binary + education + TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_3<-lm(PACC_diff ~ parous_group + time_diff + APOE_binary + education + TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])

summary(trafo_lm(mod_1))
summary(trafo_lm(mod_2))
summary(trafo_lm(mod_3))

```

no effect of WMH on PACC change between v1 and v2. No trend of A status

# Model 3: PACC ~ childbirth* wmh

```{r}


mod_1<-lm(PACC_diff ~ kids_total*volume_wmh + time_diff + APOE_binary + education +TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_2<-lm(PACC_diff ~ kids_total*volume_wmh + time_diff + APOE_binary + education+TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "parietal",])
mod_3<-lm(PACC_diff ~ kids_total*volume_wmh + time_diff + APOE_binary + education+TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])
mod_4<-lm(PACC_diff ~ kids_total*volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "occipital",])
mod_5<-lm(PACC_diff ~ kids_total*volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])


summary(trafo_lm(mod_1))
summary(trafo_lm(mod_2))
summary(trafo_lm(mod_3))
summary(trafo_lm(mod_4))
summary(trafo_lm(mod_5))

diagnostics(trafo_lm(mod_5))
plot(trafo_lm(mod_5))

p_values <- c(0.14037 ,  0.57345  , 0.54029  , 0.8180 ,  0.01995 )
#None survive, basal ganglia becomes trend

fdr_adjusted_p_values <- p.adjust(p_values)
fdr_adjusted_p_values

data_wmh_w%>%
  filter(bilateral_region == "basal ganglia")%>%
  mutate(tier = ntile(volume_wmh, 3))%>%
  ggplot(aes(x= kids_total, y= PACC_diff, colour = tier))+
  geom_jitter()+
  facet_wrap(~tier)+
  geom_smooth(method = "lm")
```
# sensitivity analysis

##CVR in original model 
```{r}
mod_5<-lm(PACC_diff ~ kids_total*volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])
mod_5_a<-lm(PACC_diff ~ kids_total*volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1+ Fuma2 + hypertension, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])

summary(trafo_lm(mod_5))
summary(trafo_lm(mod_5_a))
```
Smoking status and hypertension makes Rsquared higher 

## different domains
```{r}
#main effect

mod_1<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education  + TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_2<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "parietal",])
mod_3<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])
mod_4<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "occipital",])
mod_5<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])

summary(trafo_lm(mod_1))
summary(trafo_lm(mod_2))
summary(trafo_lm(mod_3))
summary(trafo_lm(mod_4))
summary(trafo_lm(mod_5))

tr <- trafo_lm(mod_2)

# Extract the transformed model
transformed_model <- tr$trafo_mod

# Now use plot_model()
plot_model(transformed_model, type = "pred")

#Interaction

mod_1<-lm(diff_Executive ~ kids_total*volume_wmh + time_diff + APOE_binary + education +TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_2<-lm(diff_Executive ~ kids_total*volume_wmh + time_diff + APOE_binary + education+TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "parietal",])
mod_3<-lm(diff_Executive ~ kids_total*volume_wmh + time_diff + APOE_binary + education+TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])
mod_4<-lm(diff_Executive ~ kids_total*volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "occipital",])
mod_5<-lm(diff_Executive ~ kids_total*volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])


summary(trafo_lm(mod_1))
summary(trafo_lm(mod_2))
summary(trafo_lm(mod_3))
summary(trafo_lm(mod_4))
summary(trafo_lm(mod_5))
```
Main effect:
WMH load in parietal positive relation with Executive change?? lol
Negative relation between WMH and Language in pparietal and temporal

Interaction: 
Significant in frontal and very significant in temporal lobes for executive
Significant in basal ganglia for Language

#Results

## Significant interaction of kids and wmh on PACC change in basal ganglia
```{r}
mod_pacc_bas<-lm(PACC_diff ~ kids_total*volume_wmh + A+ time_diff + APOE_binary + education+ TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "basal ganglia",])

summary(trafo_lm(mod_pacc_bas))

tr <- trafo_lm(mod_pacc_bas)

# Extract the transformed model
transformed_model <- tr$trafo_mod

# Now use plot_model()
plot_model(transformed_model, type = "int")

data_wmh_w%>%
  filter(bilateral_region == "basal ganglia")%>%
  mutate(tier = ntile(volume_wmh, 3))%>%
  ggplot(aes(x= kids_total, y= PACC_diff, colour = tier))+
  geom_jitter()+
  facet_wrap(~tier)+
  geom_smooth(method = "lm")

p_values<- c(0.14037 ,  0.57345  , 0.54029  , 0.8180 ,  0.01995 )
#None survive, basal ganglia becomes trend

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

```

##Significant positive main effect of WMH on Executive change in Parietal lobes
```{r}
mod_executive_par<-lm(diff_Executive ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1+ A, data = data_wmh_w[data_wmh_w$bilateral_region == "parietal",])

summary(trafo_lm(mod_executive_par))

diagnostics(trafo_lm(mod_executive_par))

p_values <- c(0.17234 ,  0.01078 ,0.11089 ,  0.8536  , 0.72301   )
#Becomes trend but just barely?

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

data_wmh_w%>%
  filter(bilateral_region == "parietal")%>%
  ggplot(aes(x= volume_wmh, y= PACC_diff))+
  geom_jitter()+
  geom_smooth(method = "lm")
```

##Significant negative main effect of WMH on Language in parietal and temporal
```{r}
mod_lang_par<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1 + A, data = data_wmh_w[data_wmh_w$bilateral_region == "parietal",])
mod_lang_temp<-lm(diff_Language ~ volume_wmh + time_diff + APOE_binary + education+ TIV+ age_v1 + A, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])

summary(trafo_lm(mod_lang_par))
summary(trafo_lm(mod_lang_temp))

diagnostics(trafo_lm(mod_lang_temp))

p_values <- c(0.2476 , 0.0190, 0.0118 ,0.249 , 0.5954  )
#not significant after adjusting

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values
```

##Significant interaction of kids and wmh on executive score in frontal and temporal lobes 
```{r}
mod_exe_front_int<-lm(diff_Executive ~ kids_total*volume_wmh + time_diff + APOE_binary + education +TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "frontal",])
mod_exe_temp_int<-lm(diff_Executive ~ kids_total*volume_wmh + time_diff + APOE_binary + education +TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])

summary(trafo_lm(mod_exe_front_int))
summary(trafo_lm(mod_exe_temp_int))

diagnostics(trafo_lm(mod_exe_temp_int))

p_values <- c(0.0392 ,0.100464   , 0.00183, 0.1812 ,  0.10534   )
#temporal remains very significant

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

data_wmh_w%>%
  filter(bilateral_region == "temporal")%>%
  mutate(tier = ntile(volume_wmh, 3))%>%
  ggplot(aes(x= kids_total, y= diff_Executive, colour = tier, group = tier))+
  geom_jitter()+
  facet_wrap(~tier)+
  geom_smooth(method = "lm")

mod_test<-lm(diff_Executive ~ kids_total*volume_wmh+A + time_diff + APOE_binary + education +TIV+ age_v1, data = data_wmh_w[data_wmh_w$bilateral_region == "temporal",])

summary(trafo_lm(mod_test))

tr <- trafo_lm(mod_test)

# Extract the transformed model
transformed_model <- tr$trafo_mod

# Now use plot_model()
plot_model(transformed_model, type = "int")
```
#Plots! 
##JT's aesthetics
this is needed to run the plot code (creates a function used in them)
```{r}

call_aesthethics <- function(text_size){
  th <- theme(   panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank(),
                 axis.line = element_line(linewidth = 0.5),
                 legend.position = 'right',
                 legend.text = element_text(size= text_size, family="Helvetica"),
                 text = element_text(size= text_size, family="Helvetica"),
                 strip.text.x = element_text(size = rel(0.90)),
                 strip.text.y = element_text(size = rel(0.90)),
                 axis.title.x = element_text(vjust=-0.3),
                 plot.title = element_text(hjust = 0.5, vjust = 0),
                 axis.ticks = element_line(linewidth = 0.4),
                 axis.text.x.bottom  = element_text(size = rel(0.90), margin = unit(c(t = 2.5, r = 0, b = 0, l = 0), "mm")),
                 axis.title.y = element_text(vjust = 1),
                 axis.text.y = element_text(size = rel(0.90), margin = unit(c(t = 0, r = 2.5, b = 0, l = 0), "mm")),
                 axis.ticks.length = unit(-1.2, "mm"),
                 axis.text.x.top = element_text(size = rel(0.90), margin = unit(c(t = 0, r = 0, b = 2.5, l = 0), "mm")))
  return(th)
}


th <- call_aesthethics(13)

custom_labels <- c("basal ganglia" = "Basal Ganglia", "frontal" = "Frontal Lobes", "occipital" = "Occipital Lobes", "parietal" = "Parietal Lobes", "temporal" = "Temporal Lobes")
```

## Descriptives
```{r}

```

##Mod 0
```{r}
data_wmh_w%>%
  filter(bilateral_region %in% c("temporal", "frontal", "basal ganglia"))%>%
  ggplot(aes(x= kids_total, y= volume_wmh))+
  geom_jitter(size = 3)+
  geom_smooth(method = "lm", size = 2)+
  facet_wrap(~bilateral_region, scale = "free", labeller = labeller(bilateral_region = custom_labels))+
  call_aesthethics(25)+
  labs(x = "Number of childbirths", y = "WMH volume")

ggsave("G:/My Drive/AAIC 2025/poster/mod1plot.png", width = 13.5, height = 8.5)
```

##Mod 2 - positive main effect

```{r}
data_wmh_w%>%
  filter(bilateral_region == "parietal")%>%
  ggplot(aes(x= diff_Executive, y= volume_wmh))+
  geom_jitter(size = 3)+
  geom_smooth(method = "lm", size = 2)
  
```

## Mod 3 - sig interaction

```{r}
#this one is fine but I feel like it sends the wrong message?

data_wmh_w%>%
  filter(bilateral_region %in% c("temporal"))%>%
  mutate(tier = ntile(volume_wmh, 3),
         kids_total_factor = as.factor(ifelse(kids_total >= 3, "3+", kids_total)))%>%
  ggplot(aes(x= kids_total_factor, y= diff_Executive, group = kids_total_factor, colour= kids_total_factor, fill = kids_total_factor))+
  geom_violin(alpha=0.2, size= 1.5,color = NA)+
  geom_jitter(size = 2)+
  facet_wrap(~tier)

#this one maybe better
data_wmh_w%>%
  filter(bilateral_region %in% c("temporal"))%>%
  mutate(tier = ntile(volume_wmh, 3),
         kids_total_factor = as.factor(ifelse(kids_total >= 3, "3+", kids_total)))%>%
  ggplot(aes(x= kids_total, y= diff_Executive))+
  geom_jitter(size = 2)+
  geom_line(data = preds, aes(x= x, y= predicted_bt, group=`WMH Load`, colour=`WMH Load`), inherit.aes = F, size=2)+
  geom_ribbon(data = preds, aes(x= x, ymin = conf.low_bt, ymax = conf.high_bt, fill=`WMH Load`),  alpha=.2, inherit.aes = F)+
  call_aesthethics(25)+
  labs(y = "Executive score difference (V2-V1)", x = "Number of childbirths")+
  scale_color_manual(labels = c("Low", "Mid", "High"), values = c("#345DA4", "#7D9DE5", "#AABDF3")) +
  scale_fill_manual(labels = c("Low", "Mid", "High"), values = c("#345DA4", "#7D9DE5", "#AABDF3")) +
  facet_wrap(~bilateral_region,labeller = labeller(bilateral_region = custom_labels))

data_wmh_w%>%
  filter(bilateral_region %in% c("frontal"))%>%
  mutate(tier = ntile(volume_wmh, 3),
         kids_total_factor = as.factor(ifelse(kids_total >= 3, "3+", kids_total)))%>%
  ggplot(aes(x= kids_total, y= diff_Executive))+
  geom_jitter(size = 2)+
  geom_line(data = preds2, aes(x= x, y= predicted_bt, group=`WMH Load`, colour=`WMH Load`), inherit.aes = F, size=2)+
  geom_ribbon(data = preds2, aes(x= x, ymin = conf.low_bt, ymax = conf.high_bt, fill=`WMH Load`),  alpha=.2, inherit.aes = F)+
  call_aesthethics(25)+
  labs(y = "Executive score difference (V2-V1)", x = "Number of childbirths")+
  scale_color_manual(labels = c("Low", "Mid", "High"), values = c("#345DA4", "#7D9DE5", "#AABDF3")) +
  scale_fill_manual(labels = c("Low", "Mid", "High"), values = c("#345DA4", "#7D9DE5", "#AABDF3")) +
  facet_wrap(~bilateral_region, labeller = labeller(bilateral_region = custom_labels))
  
#Creating pred values for fitting lines on plot for temporal

tr <- trafo_lm(mod_exe_temp_int)

# Extract the transformed model
transformed_model <- tr$trafo_mod
mf <- model.frame(transformed_model)
names(mf)[1] <- "diff_Executive"  # or your original outcome variable name

# Update the model frame in the model object
transformed_model$model <- mf


preds <-as_tibble(ggpredict(transformed_model, terms = c("kids_total [0:4]", "volume_wmh"), back.transform = F))%>%
  rename(`WMH Load` = group)%>%
  mutate(bilateral_region ="temporal",
         predicted_bt = ((tr$lambda * preds$predicted + 1)^(1 / tr$lambda))- 2.598893,
         conf.high_bt = ((tr$lambda * preds$conf.high + 1)^(1 / tr$lambda))- 2.598893,
         conf.low_bt = ((tr$lambda * preds$conf.low + 1)^(1 / tr$lambda))- 2.598893)

#Now for frontal
tr2 <- trafo_lm(mod_exe_front_int)

# Extract the transformed model
transformed_model2 <- tr2$trafo_mod
mf <- model.frame(transformed_model2)
names(mf)[1] <- "diff_Executive"  # or your original outcome variable name

# Update the model frame in the model object
transformed_model2$model <- mf


preds2 <-as_tibble(ggpredict(transformed_model2, terms = c("kids_total [0:4]", "volume_wmh"), back.transform = F))%>%
  rename(`WMH Load` = group)%>%
  mutate(bilateral_region ="frontal",
         predicted_bt = ((tr2$lambda * preds2$predicted + 1)^(1 / tr2$lambda))- 2.598893,
         conf.high_bt = ((tr2$lambda * preds2$conf.high + 1)^(1 / tr2$lambda))- 2.598893,
         conf.low_bt = ((tr2$lambda * preds2$conf.low + 1)^(1 / tr2$lambda))- 2.598893)

preds_all<-full_join(preds, preds2)
```






