---
title: "ADPD 2026 abstract"
author: "Clazz"
date: "2025-07-07"
output: html_document
---

```{r}
library(readxl)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("psych")
library(psych)
#install.packages("tidyverse")
library(tidyverse)
#install.packages("ggpubr")
library(ggpubr)
#install.packages("broom")
library(broom)
#install.packages("ggeffects")
library(ggeffects)
#install.packages("MASS")
library(MASS)
#install.packages("mice")
library(mice)
#install.packages("VIM")
library(VIM)
#install.packages("car")
library(car)
#install.packages("sjPlot")
library(sjPlot)
#install.packages("caret")
library(caret)
#install.packages("bestNormalize")
library(bestNormalize)
#install.packages("ggrain")
library(ggrain)
#install.packages("emmeans")
library(emmeans)
#install.packages("skimr")
library(skimr)
library(glmnet)
#install.packages("bullseye")
library(bullseye)
#install.packages("lavaan")
library(lavaan)
#install.packages("brms")
library(brms)
#install.packages("mgcv")
library(mgcv)
library(splines)
#install.packages("fitdistrplus")
library(fitdistrplus)
#install.packages("patchwork")
library(patchwork)
#install.packages("lme4")
library(lme4)
#install.packages("lmerTest")
library(lmerTest)
#install.packages("MuMIn")
library(MuMIn)
```

# Read in data

reading in basic, biomarkers and mri dates

```{r, echo=FALSE}
data<-read_delim("G:/My Drive/Data/ALFA+/ALFA_dataset_basic.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)

date_mri_v1<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_scandates_ALFA_PLUS_V1_20231120_20250210_135308.xlsx")%>%
  dplyr::select(subject_label, scandate)%>%
  mutate(subject_label = as.numeric(subject_label))%>%
  rename(subject = subject_label,
         scandate_v1 = scandate)

date_mri_v2<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_scandates_ALFA_PLUS_V2_20230518_20250210_135351.xlsx")%>%
  dplyr::select(subject_label, scandate)%>%
  mutate(subject_label = as.numeric(subject_label))%>%
  rename(subject = subject_label,
         scandate_v2 = scandate)

date_mri_long<-full_join(date_mri_v1, date_mri_v2)%>%
  mutate(scandate_v1 = as.Date(scandate_v1, format = "%Y-%m-%d"),
         scandate_v2 = as.Date(scandate_v2, format = "%Y-%m-%d"),
         diff_mri = as.numeric(difftime(scandate_v2, scandate_v1, units = c("days")))/365.25)

#adding MRI sessions dates to the dataset
data<-left_join(data, date_mri_long)%>%
  mutate(DOB = as.Date (DOB, format = "%Y-%m-%d"),
         age_v1_mri = as.numeric(difftime(scandate_v1, DOB, units= c("days")))/365.25,
         age_v2_mri = as.numeric(difftime(scandate_v2, DOB, units= c("days")))/365.25)

biomarkers<-read_delim("G:/My Drive/Data/ALFA+/biomarkers/biomarkers_CSF_processed.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)%>%
  filter(Visit == "V1")%>%
  dplyr::select(-Visit)

data<-left_join(data, biomarkers)


```

#Cognitive results - LMM

```{r, echo=FALSE}


cognition<-read_excel("G:/My Drive/Data/ALFA+/Cognition/PACCv1_to_v3(incomplete).xlsx")%>%
  dplyr::select(IdParticipante, Visit, Z_PACC, Z_FCSRTI_RT, Z_aWMS_DR_RS, Z_Wclave_totalscr, Z_FS_totalcnswr, FechaResultado)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = as.numeric(subject))

data_cog_long<- left_join(data, cognition)%>%
  mutate(age_cog = as.numeric(difftime(FechaResultado, DOB, units= c("days")))/365.25)%>%
  group_by(subject) %>%
  filter(Visit != 3,
         n() == 2) %>%
  ungroup()


fit_cog<-lmer(Z_FS_totalcnswr ~ kids_total*A +age_cog + APOE_binary + education +(1 | subject), data_cog_long[data_cog_long$sex == 2,])
summary(fit_cog)
r.squaredGLMM(fit_cog)

plot_model(fit_cog, type = "int")

###Composites

composites1<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 1 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -Z_PACC_Composite_1.2._v1 )%>%
  rename(subject = IdParticipante,
         Z_Attention_Composite = Z_Attention_Composite_v1,
         Z_Memory_Composite = Z_Memory_Composite_v1,
         Z_Executive_Composite = Z_Executive_Composite_v1,
         Z_Language_Composite = Z_Language_Composite_v1,
         Z_Visual_Composite = Z_Visual_Composite_v1)%>%
  mutate(Visit = "1")

composites2<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 2 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -V2S1_dt, -Z_PACC_Composite_1.2._V2 )%>%
  rename(subject = IdParticipante,
         Z_Attention_Composite = Z_Attention_Composite_V2,
         Z_Memory_Composite = Z_Memory_Composite_V2,
         Z_Executive_Composite = Z_Executive_Composite_V2,
         Z_Language_Composite = Z_Language_Composite_V2,
         Z_Visual_Composite = Z_Visual_Composite_V2)%>%
  mutate(Visit = "2")

composite_all<-full_join(composites1, composites2)%>%
  group_by(subject) %>%
  ungroup()%>%
  mutate(subject = as.numeric(subject))

data_cog_long<-left_join(data_cog_long, composite_all)

fit_cog_domain<-lmer(Z_Visual_Composite ~ kids_total*A +age_cog + APOE_binary + education +age_menopause+(1 | subject), data_cog_long[data_cog_long$sex == 2,])
summary(fit_cog_domain)
r.squaredGLMM(fit_cog_domain)
```

#Cognitive results - LM of change
test
test again
```{r}
composites1<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 1 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -Z_PACC_Composite_1.2._v1 )%>%
  rename(subject = IdParticipante)

composites2<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 2 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -V2S1_dt, -Z_PACC_Composite_1.2._V2 )%>%
  rename(subject = IdParticipante)

PACC_diff<-read_excel("G:/My Drive/Data/ALFA+/Cognition/PACCv1_to_v3(incomplete).xlsx")%>%
  dplyr::select(IdParticipante, Visit, Z_PACC, FechaResultado)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = as.numeric(subject))%>%
  filter(Visit != 3)%>%
  pivot_wider(names_from = Visit, values_from = c(Z_PACC, FechaResultado))%>%
  mutate(PACC_diff = Z_PACC_2 - Z_PACC_1,
         time_diff = as.numeric(difftime(FechaResultado_2, FechaResultado_1, units= c("days"))))%>%
  dplyr::select(subject, PACC_diff, time_diff)

composite_all_diff<-full_join(composites1, composites2)%>%
  mutate(diff_Attention = Z_Attention_Composite_V2 - Z_Attention_Composite_v1,
         diff_Memory = Z_Memory_Composite_V2 - Z_Memory_Composite_v1,
         diff_Executive = Z_Executive_Composite_V2 - Z_Executive_Composite_v1,
         diff_Language = Z_Language_Composite_V2 - Z_Language_Composite_v1,
         diff_Visual = Z_Visual_Composite_V2 - Z_Visual_Composite_v1,
         subject = as.numeric(subject))%>%
  dplyr:: select(subject, diff_Attention, diff_Memory, diff_Executive, diff_Language, diff_Visual)

data_cog_diff<-left_join(data, composite_all_diff)%>%
  left_join(., PACC_diff)%>%
  mutate(A=factor(A),
         APOE_binary=factor(APOE_binary),
         AT=factor(AT))

mod_pacc<-lm(PACC_diff ~ kids_total*A*sex + age_v1 + education + APOE_binary + time_diff, data = data_cog_diff)
summary(mod_pacc)

mod_pacc_w<-lm(PACC_diff ~ kids_total*A + A+age_v1 + education + APOE_binary + time_diff, data = data_cog_diff[data_cog_diff$sex==2,])
summary(mod_pacc_w)

tidy_mod1_es<-tidy(mod_pacc_w)$estimate
tidy_mod1_p<-tidy(mod_pacc_w)$p.value


hist(data_cog_diff$PACC_diff)

plot_model(mod_pacc_w, type = "int")

#trying with poly
mod_pacc_w_poly<-lm(PACC_diff ~ poly(kids_total, 2, raw = TRUE)*A+ age_menopause_pmm +age_v1 + education + APOE_binary + time_diff, data = data_cog_diff[data_cog_diff$sex==2,])
summary(mod_pacc_w_poly)
```

#hippo vol change
```{r}
hippo<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_freesurfer7_hippoSfVolumes_ALFA_PLUS_V1_20231120_20250208_111212.xlsx")%>%
  group_by(subject, region) %>%
  summarize(bilateral_volume = sum(value), .groups = "drop")%>%
 mutate(region = case_when(region == "CA1-body"| region == "CA1-head" ~ "CA1",
            region == "CA3-body"| region == "CA3-head" ~ "CA3",
            region == "CA4-body"| region == "CA4-head" ~ "CA4",
            region == "GC-ML-DG-body" | region == "GC-ML-DG-head" ~ "GC_ML_DG",
            region == "molecular_layer_HP-body" | region == "molecular_layer_HP-head"~ "molecular_layer_HP",
            region == "presubiculum-body" | region == "presubiculum-head" ~ "presubiculum",
            region == "subiculum-body" | region == "subiculum-head" ~ "subiculum",
            .default = region))%>%
  group_by(region, subject)%>%
  summarize( bilateral_volume = sum(bilateral_volume), .groups = "drop")%>%
  pivot_wider(names_from = region, values_from = bilateral_volume, names_prefix = "vol_")%>%
  mutate(subject = as.numeric(subject),
         Visit = "V1")

hippo_data<-left_join(data, hippo)

hippo2<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_freesurfer7_hippoSfVolumes_ALFA_PLUS_V2_20230518_20250208_112459.xlsx")%>%
  group_by(subject, region) %>%
  summarize(bilateral_volume = sum(value), .groups = "drop")%>%
  mutate(region = case_when(region == "CA1-body"| region == "CA1-head" ~ "CA1",
            region == "CA3-body"| region == "CA3-head" ~ "CA3",
            region == "CA4-body"| region == "CA4-head" ~ "CA4",
            region == "GC-ML-DG-body" | region == "GC-ML-DG-head" ~ "GC_ML_DG",
            region == "molecular_layer_HP-body" | region == "molecular_layer_HP-head"~ "molecular_layer_HP",
            region == "presubiculum-body" | region == "presubiculum-head" ~ "presubiculum",
            region == "subiculum-body" | region == "subiculum-head" ~ "subiculum",
            .default = region))%>%
  group_by(region, subject)%>%
  summarize( bilateral_volume = sum(bilateral_volume), .groups = "drop")%>%
  pivot_wider(names_from = region, values_from = bilateral_volume, names_prefix = "vol_")%>%
  mutate(subject = as.numeric(subject),
         Visit = "V2")

hippo_data2<-left_join(data, hippo2)
  

hippo_data_long<-bind_rows(hippo_data, hippo_data2)%>%
  mutate(Visit = factor(Visit))%>%
  mutate(age_mri = case_when(
    Visit == "V1" ~ age_v1_mri,
    Visit == "V2" ~ age_v2_mri
  ))%>%
  dplyr::select(-age_v1_mri, -age_v2_mri)%>%
  drop_na(scandate_v2)

#fucntion to run 
run_volume_lmer_regressions <- function(data, predictors) {
  # Find all columns that start with "vol_"
  outcome_vars <- grep("^vol_", names(data), value = TRUE)

  # Create a formula string from predictors
  predictor_formula <- paste(predictors, collapse = " + ")

  # Run regressions and store summary results
  summaries <- lapply(outcome_vars, function(outcome) {
    formula <- as.formula(paste0("`", outcome, "` ~ ", predictor_formula))
    model <- lmer(formula, data = data)
    summary(model)
  })

  # Name each summary by its outcome variable
  names(summaries) <- outcome_vars

  return(summaries)
}

##no interaction

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total", "age_mri", "APOE_binary", "scale(TIV)", "(1 | subject)"))

p_values <- c(0.0915 ,0.551245  ,  0.5645  ,  0.4297   , 0.0825, 0.83811 ,   0.0025, 0.1691  ,  0.934 ,   0.548,    0.455    )

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

#fimbria significant, fimbria survives

##interaction

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*A", "age_mri", "APOE_binary", "scale(TIV)", "(1 | subject)"))

p_values <- c(0.6607 ,   0.14130   , 0.0158, 0.0326, 0.718  ,  0.0965, 0.8346  ,0.582  ,  0.00403, 0.0139 ,0.0253  )

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

#CA4 0.0158 , GC 0.0326, parasubiculum 0.00403, presubiculum 0.0139, subiculum 0.0253 sig
#parasubiculum survives, trend for the others

mod_test<-lmer(vol_parasubiculum ~ kids_total*A + age_mri+ APOE_binary+ scale(TIV)+ (1 | subject), data = hippo_data_long[hippo_data_long$sex==2,])
summary(mod_test)
plot_model(mod_test, type="int")
r.squaredGLMM(mod_test)

```

# plots

## Plot aesthetics
this is needed to run the plot code (creates a function used in them)
```{r}

call_aesthethics <- function(text_size){
  th <- theme(   panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank(),
                 axis.line = element_line(linewidth = 0.5),
                 legend.position = 'right',
                 legend.text = element_text(size= text_size, family="Helvetica"),
                 text = element_text(size= text_size, family="Helvetica"),
                 strip.text.x = element_text(size = rel(0.90)),
                 strip.text.y = element_text(size = rel(0.90)),
                 axis.title.x = element_text(vjust=-0.3),
                 plot.title = element_text(hjust = 0.5, vjust = 0),
                 axis.ticks = element_line(linewidth = 0.4),
                 axis.text.x.bottom  = element_text(size = rel(0.90), margin = unit(c(t = 2.5, r = 0, b = 0, l = 0), "mm")),
                 axis.title.y = element_text(vjust = 1),
                 axis.text.y = element_text(size = rel(0.90), margin = unit(c(t = 0, r = 2.5, b = 0, l = 0), "mm")),
                 axis.ticks.length = unit(-1.2, "mm"),
                 axis.text.x.top = element_text(size = rel(0.90), margin = unit(c(t = 0, r = 0, b = 2.5, l = 0), "mm")))
  return(th)
}


th <- call_aesthethics(13)

```

## PACC ~ childbirth*AB

```{r}
data_cog_diff%>%
  filter(sex == 2)%>%
  drop_na(A)%>%
  ggplot(aes(x=kids_total, y=PACC_diff, group=A, colour=A))+
  geom_jitter()+
  geom_smooth(method = "lm")+
  call_aesthethics(15)+
  labs(x = "Number of childbirths", y = "PACC change (Z-score)")
```

## hipp ~ childbirth*AB

### plotting both timepoints
```{r}
hippo_data_plot_long%>%
  filter(sex== 2)%>%
  drop_na(A)%>%
  
```



### plotting differences
```{r}
#I'm aware how clunky this code is... I was fighting with pivots for a while and getting nowhere. So this is the best I could do for now sorry! 

hippo_data_plot<-hippo
colnames(hippo_data_plot) <- paste(colnames(hippo_data_plot), 'V1', sep = '_')
hippo_data_plot<-hippo_data_plot%>%
  rename(subject = subject_V1)%>%
  dplyr::select(-Visit_V1)

hippo_data_plot2<-hippo2
colnames(hippo_data_plot2) <- paste(colnames(hippo_data_plot2), 'V2', sep = '_')
hippo_data_plot2<-hippo_data_plot2%>%
  rename(subject = subject_V2)%>%
  dplyr::select(-Visit_V2)

hippo_data_plot<-left_join(hippo_data_plot, hippo_data_plot2)

base_names <- names(hippo_data_plot) %>%
  str_subset("^vol_.*_V1$") %>%
  str_remove("_V1$")


for (base in base_names) {
  v1_col <- paste0(base, "_V1")
  v2_col <- paste0(base, "_V2")
  diff_col <- paste0(base, "_diff")
  
  hippo_data_plot[[diff_col]] <- hippo_data_plot[[v2_col]] - hippo_data_plot[[v1_col]]
}

hippo_data_plot<-hippo_data_plot%>%
  dplyr::select(subject, vol_CA1_diff, vol_CA3_diff, vol_CA4_diff, vol_GC_ML_DG_diff, vol_HATA_diff, vol_Hippocampal_tail_diff, vol_Whole_hippocampal_body_diff, vol_Whole_hippocampal_head_diff, vol_Whole_hippocampus_diff, vol_fimbria_diff, vol_molecular_layer_HP_diff, vol_parasubiculum_diff, vol_presubiculum_diff, vol_subiculum_diff)

hippo_data_plot<-left_join(data, hippo_data_plot)

hippo_data_plot%>%
  filter(sex==2)%>%
  mutate(A=factor(A))%>%
  drop_na(A)%>%
ggplot(aes(x=kids_total, y=vol_parasubiculum_diff, group=A, colour=A))+
  geom_jitter()+
  geom_smooth(method = "lm")+
  call_aesthethics(15)+
  labs(x = "Number of childbirths", y = "Volume change")
```

#Abstract

Objectives: Little is known about the impact of pregnancy on Alzheimer's Disease (AD) trajectory. We examined the impact of childbriths and AD pathlogy on cognitive decline, and related it to hippocampal volume change in cognitively unimpaired (CU) post-menopausal women at risk of AD.

```{r, echo=F}

pt_number<-data%>%
  nrow()%>%
  as.numeric()
  
pt_number_women<-data%>%
  filter(sex==2)%>%
  nrow()%>%
  as.numeric()

```

Methods: Our sample includes `r pt_number_women` CU post-menopausal women from the ALFA+ study. Hippocampus subfield segmentation was obtained using Freesurfer 7. Amyloid (Aβ) positivity was defined as CSF Aβ42/40 ratio\<0.071. Cognition was measured at visit 1 and 2 using PACC scores. We first examined the impact of childbirths on cognitive change, and moderation by Aβ positivity. We then assessed the effect of childbirth on hippocampal subfield volume change, and moderation by Aβ positivity. All linear models included APOE-ε4 carriership, age, time difference between visits and total intracranial volume as covariates (Table 1). FDR correction was used in models 3 and 4.

Results: A trending interaction between number of childbirths and amyloid positivity on cognitive change was found (F=`r tidy_mod1_es[8]`, p=`r tidy_mod1_p[8]`), where Aβ+ women showed worse decline with higher parity. Childbirths had a significant positive main effect on volume change in the fimbria (F=5.240, adj.p.=0.020). A significant interaction of childbirths and Aβ+ was found in the parasubiculum (F=6.8155, adj.p.=0.0443), such that Aβ+ women showed worse volume reduction with higher parity. This interaction was trending in the CA4, dentate gyrus, presubiculum and subiculum.

Conclusion: In CU post-menopausal women at risk of AD, parity might moderate the impact of AD pathology on cognitive change at 3 years. Additionally, parity independently influences volume change in the fimbria, and interacts with AD pathology on parasubiculum volume. Results highlight the effect of parity on AD-related neuropathology via hippocampal volume change. 

