---
title: "ADPD 2026 abstract"
author: "Clazz"
date: "2025-07-07"
output: html_document
---

```{r}
library(readxl)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("psych")
library(psych)
#install.packages("tidyverse")
library(tidyverse)
#install.packages("ggpubr")
library(ggpubr)
#install.packages("broom")
library(broom)
#install.packages("ggeffects")
library(ggeffects)
#install.packages("MASS")
library(MASS)
#install.packages("mice")
library(mice)
#install.packages("VIM")
library(VIM)
#install.packages("car")
library(car)
#install.packages("sjPlot")
library(sjPlot)
#install.packages("caret")
library(caret)
#install.packages("bestNormalize")
library(bestNormalize)
#install.packages("ggrain")
library(ggrain)
#install.packages("emmeans")
library(emmeans)
#install.packages("skimr")
library(skimr)
library(glmnet)
#install.packages("bullseye")
library(bullseye)
#install.packages("lavaan")
library(lavaan)
#install.packages("brms")
library(brms)
#install.packages("mgcv")
library(mgcv)
library(splines)
#install.packages("fitdistrplus")
library(fitdistrplus)
#install.packages("patchwork")
library(patchwork)
#install.packages("lme4")
library(lme4)
#install.packages("lmerTest")
library(lmerTest)
#install.packages("MuMIn")
library(MuMIn)
```

# Read in data

reading in basic, biomarkers and mri dates

```{r, echo=FALSE}
data<-read_delim("G:/My Drive/Data/ALFA+/ALFA_dataset_basic.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)

date_mri_v1<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_scandates_ALFA_PLUS_V1_20231120_20250210_135308.xlsx")%>%
  dplyr::select(subject_label, scandate)%>%
  mutate(subject_label = as.numeric(subject_label))%>%
  rename(subject = subject_label,
         scandate_v1 = scandate)

date_mri_v2<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_scandates_ALFA_PLUS_V2_20230518_20250210_135351.xlsx")%>%
  dplyr::select(subject_label, scandate)%>%
  mutate(subject_label = as.numeric(subject_label))%>%
  rename(subject = subject_label,
         scandate_v2 = scandate)

date_mri_long<-full_join(date_mri_v1, date_mri_v2)%>%
  mutate(scandate_v1 = as.Date(scandate_v1, format = "%Y-%m-%d"),
         scandate_v2 = as.Date(scandate_v2, format = "%Y-%m-%d"),
         diff_mri = as.numeric(difftime(scandate_v2, scandate_v1, units = c("days")))/365.25)

#adding MRI sessions dates to the dataset
data<-left_join(data, date_mri_long)%>%
  mutate(DOB = as.Date (DOB, format = "%Y-%m-%d"),
         age_v1_mri = as.numeric(difftime(scandate_v1, DOB, units= c("days")))/365.25,
         age_v2_mri = as.numeric(difftime(scandate_v2, DOB, units= c("days")))/365.25)

biomarkers<-read_delim("G:/My Drive/Data/ALFA+/biomarkers/biomarkers_CSF_processed.csv", 
    delim = ",", escape_double = FALSE, trim_ws = TRUE)%>%
  filter(Visit == "V1")%>%
  dplyr::select(-Visit)

data<-left_join(data, biomarkers)%>%
  mutate(A=factor(A),
         APOE_binary=factor(APOE_binary),
         HRT = factor(HRT))


```

# Cognitive results - LMM
without time interaction executive scores are significant
with time interaction the pacc scores are
## PACC
```{r, echo=FALSE}
#frist getting the time difference beteen cog tests
time_diff<-read_excel("G:/My Drive/Data/ALFA+/Cognition/PACCv1_to_v3(incomplete).xlsx")%>%
  dplyr::select(IdParticipante, Visit,FechaResultado)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = as.numeric(subject))%>%
  filter(Visit != 3)%>%
  pivot_wider(names_from = Visit, values_from = FechaResultado, names_prefix = c("Fecha_Resultado"))%>%
  mutate(time_diff = as.numeric(difftime(Fecha_Resultado2, Fecha_Resultado1, units= c("days"))))%>%
  dplyr::select(subject, time_diff)

cognition<-read_excel("G:/My Drive/Data/ALFA+/Cognition/PACCv1_to_v3(incomplete).xlsx")%>%
  dplyr::select(IdParticipante, Visit, Z_PACC, Z_FCSRTI_RT, Z_aWMS_DR_RS, Z_Wclave_totalscr, Z_FS_totalcnswr, FechaResultado)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = as.numeric(subject))

data_cog_long<- left_join(data, cognition)%>%
  mutate(age_cog = as.numeric(difftime(FechaResultado, DOB, units= c("days")))/365.25)%>%
  group_by(subject)%>%
  filter(Visit != 3) %>%
  filter(n()>=2)%>%
  relocate(Visit, .after=sex)%>%
  ungroup()%>%
  left_join(., time_diff)%>%
  mutate(time_diff = ifelse(Visit == 1, 0, time_diff))

#cross-sectional

fit_cog<-lmer(Z_PACC ~ kids_total*CSF_Abeta_ratio +age_cog + APOE_binary + education +(1 | subject), data_cog_long[data_cog_long$sex == 2,])
summary(fit_cog)

#both sexes
fit_cog_sex<-lmer(Z_PACC ~ kids_total*A*sex +age_cog + APOE_binary + education +(1 | subject), data_cog_long)
summary(fit_cog_sex)
```


```{r, echo=FALSE}
#Longitudinal

fit_cog<-lmer(Z_PACC ~ kids_total*A*time_diff +age_v1 + APOE_binary + education +(1 | subject), data_cog_long[data_cog_long$sex == 2,])
summary(fit_cog)
r.squaredGLMM(fit_cog)

plot_model(fit_cog, type = "int")
```

## Composites
```{r, echo=FALSE}

composites1<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 1 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -Z_PACC_Composite_1.2._v1 )%>%
  rename(subject = IdParticipante,
         Z_Attention_Composite = Z_Attention_Composite_v1,
         Z_Memory_Composite = Z_Memory_Composite_v1,
         Z_Executive_Composite = Z_Executive_Composite_v1,
         Z_Language_Composite = Z_Language_Composite_v1,
         Z_Visual_Composite = Z_Visual_Composite_v1)%>%
  mutate(Visit = "1")

composites2<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 2 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -V2S1_dt, -Z_PACC_Composite_1.2._V2 )%>%
  rename(subject = IdParticipante,
         Z_Attention_Composite = Z_Attention_Composite_V2,
         Z_Memory_Composite = Z_Memory_Composite_V2,
         Z_Executive_Composite = Z_Executive_Composite_V2,
         Z_Language_Composite = Z_Language_Composite_V2,
         Z_Visual_Composite = Z_Visual_Composite_V2)%>%
  mutate(Visit = "2")

composite_all<-full_join(composites1, composites2)%>%
  group_by(subject) %>%
  ungroup()%>%
  mutate(subject = as.numeric(subject))

data_cog_long_composites<-left_join(data_cog_long, composite_all)%>%
  mutate(A= factor(A),
         APOE_binary = factor(APOE_binary))

#Cross sectional

fit_cog_domain<-lmer(Z_Executive_Composite ~ kids_total*CSF_pTau217 +age_cog + APOE_binary + education + (1 | subject), data_cog_long_composites[data_cog_long_composites$sex == 2,])
summary(fit_cog_domain)


p_values <- c(0.33549      , 0.7260  ,   0.02337 , 0.36791     ,0.1997    )

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

plot_model(fit_cog_domain, type= "int")

#both sexes

fit_cog_domain_sex<-lmer(Z_Visual_Composite ~ kids_total*A*sex +age_cog + APOE_binary + education + (1 | subject), data_cog_long_composites)
summary(fit_cog_domain)
```


```{r, echo=FALSE}
# Longitudinal

fit_cog_domain_long<-lmer(Z_Attention_Composite ~ kids_total*CSF_pTau217*time_diff +age_v1 + APOE_binary + education + (1 | subject), data_cog_long_composites[data_cog_long_composites$sex == 2,])
summary(fit_cog_domain_long)

#not significant with A status or ratio, significant with pTau217

```



# hippo vol change

```{r}
hippo<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_freesurfer7_hippoSfVolumes_ALFA_PLUS_V1_20231120_20250208_111212.xlsx")%>%
  group_by(subject, region) %>%
  summarize(bilateral_volume = sum(value), .groups = "drop")%>%
 mutate(region = case_when(region == "CA1-body"| region == "CA1-head" ~ "CA1",
            region == "CA3-body"| region == "CA3-head" ~ "CA3",
            region == "CA4-body"| region == "CA4-head" ~ "CA4",
            region == "GC-ML-DG-body" | region == "GC-ML-DG-head" ~ "GC_ML_DG",
            region == "molecular_layer_HP-body" | region == "molecular_layer_HP-head"~ "molecular_layer_HP",
            region == "presubiculum-body" | region == "presubiculum-head" ~ "presubiculum",
            region == "subiculum-body" | region == "subiculum-head" ~ "subiculum",
            .default = region))%>%
  group_by(region, subject)%>%
  summarize( bilateral_volume = sum(bilateral_volume), .groups = "drop")%>%
  pivot_wider(names_from = region, values_from = bilateral_volume, names_prefix = "vol_")%>%
  mutate(subject = as.numeric(subject),
         Visit = "V1")

hippo_data<-left_join(data, hippo)

hippo2<-read_excel("G:/My Drive/Data/ALFA+/neuroimaging/bx_freesurfer7_hippoSfVolumes_ALFA_PLUS_V2_20230518_20250208_112459.xlsx")%>%
  group_by(subject, region) %>%
  summarize(bilateral_volume = sum(value), .groups = "drop")%>%
  mutate(region = case_when(region == "CA1-body"| region == "CA1-head" ~ "CA1",
            region == "CA3-body"| region == "CA3-head" ~ "CA3",
            region == "CA4-body"| region == "CA4-head" ~ "CA4",
            region == "GC-ML-DG-body" | region == "GC-ML-DG-head" ~ "GC_ML_DG",
            region == "molecular_layer_HP-body" | region == "molecular_layer_HP-head"~ "molecular_layer_HP",
            region == "presubiculum-body" | region == "presubiculum-head" ~ "presubiculum",
            region == "subiculum-body" | region == "subiculum-head" ~ "subiculum",
            .default = region))%>%
  group_by(region, subject)%>%
  summarize( bilateral_volume = sum(bilateral_volume), .groups = "drop")%>%
  pivot_wider(names_from = region, values_from = bilateral_volume, names_prefix = "vol_")%>%
  mutate(subject = as.numeric(subject),
         Visit = "V2")

hippo_data2<-left_join(data, hippo2)
  

hippo_data_long<-bind_rows(hippo_data, hippo_data2)%>%
  mutate(Visit = factor(Visit))%>%
  mutate(age_mri = case_when(
    Visit == "V1" ~ age_v1_mri,
    Visit == "V2" ~ age_v2_mri
  ))%>%
  dplyr::select(-age_v1_mri, -age_v2_mri)%>%
  drop_na(scandate_v2)%>%
  mutate(diff_mri = ifelse(Visit == "V1", 0, diff_mri))

#fucntion to run 
run_volume_lmer_regressions <- function(data, predictors) {
  # Find all columns that start with "vol_"
  outcome_vars <- grep("^vol_", names(data), value = TRUE)

  # Create a formula string from predictors
  predictor_formula <- paste(predictors, collapse = " + ")

  # Run regressions and store summary results
  summaries <- lapply(outcome_vars, function(outcome) {
    formula <- as.formula(paste0("`", outcome, "` ~ ", predictor_formula))
    model <- lmer(formula, data = data)
    summary(model)
  })

  # Name each summary by its outcome variable
  names(summaries) <- outcome_vars

  return(summaries)
}

##no interaction
#introduced A as a covariate to test if number of kids is truly independent

mod_simple<-lmer(vol_Whole_hippocampal_body ~ kids_total*diff_mri + age_v1 + APOE_binary + A + scale(TIV) + (1|subject), data = hippo_data_long[hippo_data_long$sex==2,])
summary(mod_simple)

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*diff_mri", "age_v1"  ,"APOE_binary","A", "scale(TIV)", "(1 | subject)"))

p_values <- c(0.2839  ,   0.67670   , 0.604591    ,0.4934 ,   0.086665 ,0.697483   , 0.022171 ,0.7787   , 0.87069 ,   0.408 ,   0.507719     )

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

#fimbria significant but does not survive

##interaction

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*A*diff_mri" ,"age_v1", "APOE_binary", "scale(TIV)", "(1 | subject)"))

p_values <- c(0.65234  ,  0.084370, 0.015109, 0.0284 ,0.413253  ,  0.105757   , 0.927819    ,0.6756  ,  0.00188, 0.00926 ,0.021231 )

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

#CA4, GC-ML-DG, subiculum and presubiculum near significant interaction between kids and A; sig for parasubiculum. no 3 way interaction with time, or effect of time.

#need to check the effect per A group; then can adjust 
emmeans(mod_1, ~ kids_total*A*diff_mri, data=hippo_data_long[hippo_data_long$sex ==2,], at=list(kids_total=c(0:4)))%>%
test()

#sensitivity analysis female var

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*A*diff_mri" ,"age_mri", "APOE_binary", "age_menopause_pmm","scale(TIV)", "(1 | subject)"))

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*A*diff_mri" ,"age_mri", "APOE_binary", "repro_span","scale(TIV)", "(1 | subject)"))

p_values <- c(0.9118  ,  0.2414  ,  0.0267 ,0.0533, 0.703, 0.2047  ,  0.8419  ,  0.759 ,   0.00234, 0.0117 ,0.0224 )

fdr_adjusted_p_values <- p.adjust(p_values, method = "fdr")
fdr_adjusted_p_values

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*A*diff_mri" ,"age_mri", "APOE_binary", "repro_span","HRT","scale(TIV)", "(1 | subject)"))

```

##longitudinal analysis
results null after introducing 3 way interaction. 
```{r}
##no interaction

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*diff_mri", "age_v1", "APOE_binary", "scale(TIV)", "(1 | subject)"))

##interaction

hippo_data_long%>%
  filter(sex == 2 )%>%
  dplyr::select(-vol_Whole_hippocampal_body, -vol_Whole_hippocampal_head, -vol_Whole_hippocampus, -`vol_hippocampal-fissure`)%>%
run_volume_lmer_regressions( .,c( "kids_total*CSF_pTau217*diff_mri" ,"age_v1", "APOE_binary", "scale(TIV)", "repro_span", "(1 | subject)"))

```

# Jack
```{r}
jack<-read_excel("G:/My Drive/Data/ALFA+/Jack signature ALFA+/bx_signature_jack_ALFA_PLUS_V1_20231120_20250527_142751.xlsx")%>%
  filter(weighted == "TRUE")%>%
  dplyr::select(-ID, -signature, -weighted)%>%
  pivot_wider(names_from = measurement, values_from = value, names_prefix = "jack_")%>%
  mutate(subject = as.numeric(subject),
         Visit = "V1")

jack2<-read_excel("G:/My Drive/Data/ALFA+/Jack signature ALFA+/bx_signature_jack_ALFA_PLUS_V2_20230518_20250210_114442.xlsx")%>%
  filter(weighted == "TRUE")%>%
  dplyr::select(-ID, -signature, -weighted)%>%
  pivot_wider(names_from = measurement, values_from = value, names_prefix = "jack_")%>%
  mutate(subject = as.numeric(subject),
         Visit = "V2")

jack_data<-bind_rows(jack, jack2)
diff_mri_df<-hippo_data_long%>%
  dplyr::select(subject, diff_mri, Visit)

jack_data_long<-left_join(hippo_data_long, jack_data)%>%
  left_join(., diff_mri_df)
  

#women
fit_jack_cort_w <- lmer(jack_ThickAvg ~ kids_total*CSF_pTau217*diff_mri  +age_v1 + APOE_binary + scale(TIV) + (1 | subject) , jack_data_long[jack_data_long$sex==2,]) 
summary(fit_jack_cort_w)

fit_jack_grey_w <- lmer(jack_GrayVol ~ kids_total*CSF_pTau217  +age_mri + APOE_binary + scale(TIV) + (1 | subject) , jack_data_long[jack_data_long$sex==2,]) 
summary(fit_jack_grey_w)

plot_model(fit_jack_grey_w, type= "int")
```


#Mediation test
does not yield much
should I test mediation with biomarker? Depends if enough proof that parity might influence AD biomarkers directly, which I doubt
```{r}

cols<-data_cog_long_composites%>%
  dplyr::select(subject, Visit, Z_Attention_Composite, Z_Executive_Composite, Z_Memory_Composite, Z_Language_Composite, Z_Visual_Composite, age_cog)

mediation_data<-hippo_data_long%>%
  mutate(Visit = ifelse(Visit == "V2", "2", "1"))%>%
  left_join(., cols)%>%
  drop_na(Z_Language_Composite)%>%
  drop_na(vol_parasubiculum)

#library(boot)

# Fit the models
med_model <- lmer(vol_parasubiculum ~ kids_total * A + age_mri + APOE_binary + TIV  +(1 | subject), 
                  data = mediation_data)

out_model <- lmer(Z_Memory_Composite ~ vol_parasubiculum + kids_total * A + age_mri + APOE_binary + TIV  +(1 | subject), 
                  data = mediation_data)

# Extract coefficients for indirect effect
a_coef <- fixef(med_model)["kids_total:A"]  # interaction effect on mediator
b_coef <- fixef(out_model)["vol_parasubiculum"]  # mediator effect on outcome

# Point estimate of indirect effect
indirect_effect <- a_coef * b_coef
print(paste("Indirect effect:", indirect_effect))

# For confidence intervals, use parametric bootstrap
bootstrap_indirect <- function(n_sims = 1000) {
  # Get variance-covariance matrices
  vcov_med <- vcov(med_model)
  vcov_out <- vcov(out_model)
  
  # Simulate coefficients
  a_sims <- rnorm(n_sims, a_coef, sqrt(vcov_med["kids_total:A", "kids_total:A"]))
  b_sims <- rnorm(n_sims, b_coef, sqrt(vcov_out["vol_parasubiculum", "vol_parasubiculum"]))
  
  # Calculate indirect effects
  indirect_sims <- a_sims * b_sims
  
  return(indirect_sims)
}

# Get bootstrap distribution
indirect_boot <- bootstrap_indirect(1000)

# Calculate confidence intervals
ci_lower <- quantile(indirect_boot, 0.025)
ci_upper <- quantile(indirect_boot, 0.975)

print(paste("95% CI: [", round(ci_lower, 4), ", ", round(ci_upper, 4), "]"))

## Test 
#install.packages("RMediation")
library(RMediation)

# Extract coefficients and standard errors
a_coef <- fixef(med_model)["kids_total:A"]
a_se <- sqrt(vcov(med_model)["kids_total:A", "kids_total:A"])

b_coef <- fixef(out_model)["vol_parasubiculum"] 
b_se <- sqrt(vcov(out_model)["vol_parasubiculum", "vol_parasubiculum"])

# Test mediation
mediation_test <- medci(mu.x = a_coef, mu.y = b_coef, 
                        se.x = a_se, se.y = b_se,
                        alpha = 0.05, type = "prodclin")

print(mediation_test)

```


# plots

## Plot aesthetics
this is needed to run the plot code (creates a function used in them)
```{r}

call_aesthethics <- function(text_size){
  th <- theme(   panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank(),
                 axis.line = element_line(linewidth = 0.5),
                 legend.position = 'right',
                 legend.text = element_text(size= text_size, family="Helvetica"),
                 text = element_text(size= text_size, family="Helvetica"),
                 strip.text.x = element_text(size = rel(0.90)),
                 strip.text.y = element_text(size = rel(0.90)),
                 axis.title.x = element_text(vjust=-0.3),
                 plot.title = element_text(hjust = 0.5, vjust = 0),
                 axis.ticks = element_line(linewidth = 0.4),
                 axis.text.x.bottom  = element_text(size = rel(0.90), margin = unit(c(t = 2.5, r = 0, b = 0, l = 0), "mm")),
                 axis.title.y = element_text(vjust = 1),
                 axis.text.y = element_text(size = rel(0.90), margin = unit(c(t = 0, r = 2.5, b = 0, l = 0), "mm")),
                 axis.ticks.length = unit(-1.2, "mm"),
                 axis.text.x.top = element_text(size = rel(0.90), margin = unit(c(t = 0, r = 0, b = 2.5, l = 0), "mm")))
  return(th)
}


th <- call_aesthethics(15)

```



## Executive ~ parity * A

```{r}
data_cog_long_composites%>%
  filter(sex==2)%>%
  drop_na(A)%>%
  mutate(kids_total_factor=ifelse(kids_total>=3, "3+", kids_total))%>%
  mutate(A = factor(A, levels = c("0", "1"), labels = c("Aβ-", "Aβ+")))%>%
  ggplot(aes(x= Visit, y = Z_Executive_Composite, group = interaction(Visit, kids_total_factor), color= kids_total_factor, fill= kids_total_factor))+
  geom_point(alpha=.6, position = position_dodge(width = 1))+
  geom_boxplot(position = position_dodge(width = 0.9), 
               alpha=.5)+
  facet_wrap(~A)+
  scale_color_manual(values = c("0" = "#8AD9B5", "1" = "#50B88E", "2" = "#047A56", "3+" = "#0A5940")) +
  scale_fill_manual(values = c("0" = "#8AD9B5", "1" = "#50B88E", "2" = "#047A56", "3+" = "#0A5940"))+
   th +
  labs(x = "Visit", 
       y = "Executive composite Z-score", 
       color = "Number of childbirths", 
       fill= "Number of childbirths")


ggsave("G:/My Drive/ADPD 2026/plot_executive.png", width = 9, height = 5)
```



## hipp ~ childbirth*AB

### Boxplot for parasubiculum subfield values per A, parity and visit 

```{r}
mod_plot<-lmer(vol_parasubiculum ~ kids_total*A + Visit + as.factor(APOE_binary) + scale(TIV) +(1| subject), data=hippo_data_long[hippo_data_long$sex==2,] )
summary(mod_plot)

plot_model(mod_plot, type= "est", title = "", axis.title = "Parasubiculum volume change model estimate" )

preds <- ggpredict(mod_plot, terms = c("kids_total", "A", "Visit")) 

hippo_data_long %>%
  filter(sex == 2) %>%
  drop_na(A) %>%
  mutate(kids_total_factor=ifelse(kids_total>=3, "3+", kids_total))%>%
  mutate(A = factor(A, labels = c("Aβ-", "Aβ+"))) %>%
  ggplot(aes(x = Visit, y = vol_parasubiculum, color = kids_total_factor, fill=kids_total_factor)) +
  geom_point(alpha = 0.6, 
             position = position_dodge(width = 1)) +
  geom_boxplot(position = position_dodge(width = 0.9), 
               alpha=.5)+
  facet_wrap(~A) +
  th +
  labs(x = "Visit", 
       y = "Parasubiculum volume", 
       color = "Number of childbirths", 
       fill= "Number of childbirths")+
  scale_color_manual(values = c("0" = "lightblue", "1" = "steelblue", "2" = "darkblue", "3+" = "navy", "4" = "midnightblue")) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "steelblue", "2" = "darkblue", "3+" = "navy", "4" = "midnightblue"))+
  scale_x_discrete(expand = expansion(mult = c(0.8, 0.8)))

ggsave("G:/My Drive/ADPD 2026/plot_parasubiculum_interaction.png", width = 9, height = 5)
```

### Individual model forest plot
```{r}
plot_model(mod_plot, 
           type = "est", 
           title=" ", 
           axis.labels = c("Childbirths*Aβ+", "TIV", "APOE-ε4 carrier", "Visit", "Aβ+", "Number of childbirths")) + 
  th+
  labs(y = "Parasubiculum volume change model estimate")+
  geom_pointrange(aes(y = estimate, 
                      ymin = conf.low, 
                      ymax = conf.high),
                  size = 0.8)
 
ggsave("G:/My Drive/ADPD 2026/plot_parasubiculum_estimation.png", width = 8, height = 5)
```

### Plot all coefficients for all regressions
```{r}
data_forest_plot<-hippo_data_long%>%
  filter(sex == 2 )

models<-list(
  CA1 = lmer(vol_CA1 ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  CA3 = lmer(vol_CA3 ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  `CA4` = lmer(vol_CA4 ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  `GC-ML-DG` = lmer(vol_GC_ML_DG ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  HATA = lmer(vol_HATA ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  Tail = lmer(vol_Hippocampal_tail ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  Fimbria = lmer(vol_fimbria ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  `Molecular layer` = lmer(vol_molecular_layer_HP ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
 `Parasubiculum` = lmer(vol_parasubiculum ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  `Presubiculum` = lmer(vol_presubiculum ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot),
  `Subiculum` = lmer(vol_subiculum ~ kids_total*A + scale(TIV) + age_mri + APOE_binary + (1 | subject),data = data_forest_plot)
)

# Extract coefficients and CIs
#install.packages("broom.mixed")
library(broom.mixed)
library(purrr)

pred_hippo_all <- map_dfr(models, ~ {
  broom.mixed::tidy(.x, terms = c("kids_total", "A"))   # Add consistent term column
}, .id = "region")%>%
  filter(term %in% c("kids_total", "A", "kids_total:A"))%>%
  dplyr::select(region, term, estimate, std.error)%>%
  pivot_wider(names_from = term, values_from = c(estimate, std.error))%>%
  mutate(plot_estimate_kids_A0 = estimate_kids_total,
         plot_estimate_kids_A1 = estimate_kids_total + `estimate_kids_total:A`,
         plot_conf.low_kids_A0 = estimate_kids_total - 1.96 * std.error_kids_total,
         plot_conf.high_kids_A0 = estimate_kids_total + 1.96 * std.error_kids_total,
         plot_conf.low_kids_A1 = (estimate_kids_total + `estimate_kids_total:A` - 1.96 * `std.error_kids_total:A`),
         plot_conf.high_kids_A1 = (estimate_kids_total + `estimate_kids_total:A` + 1.96 * `std.error_kids_total:A`))%>%
  dplyr::select(-estimate_kids_total, -estimate_A,-`estimate_kids_total:A` ,-`std.error_kids_total`, -std.error_A ,         -`std.error_kids_total:A`)%>%
  pivot_longer(
    cols = starts_with("plot_"),  # Keep 'term' as-is
    names_to = c(".value", "A"),  # .value creates columns like estimate, conf.low, etc.
    names_pattern = "plot_(.*)_kids_A(\\d+)"
  )%>%
  mutate(colour_group = case_when(region == "CA4" ~ "trend",
                                  region == "GC-ML-DG" ~ "trend",
                                  region == "Presubiculum" ~ "trend",
                                  region == "Subiculum" ~ "trend",
                                  region == "Parasubiculum" ~ "significant",
                                  .default = "normal"))



pred_hippo_all %>%
  mutate(A = factor(A, levels = c("1", "0"), labels = c("Positive", "Negative"))) %>%
  ggplot(aes(x = estimate, y = region, group = A, shape = A, color = colour_group)) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), 
                  position = position_dodge(width = 0.8),
                  size = 0.6) +
  labs(x = "Estimated effect of parity", 
       y = "Hippocampal subfield", 
       shape = "Aβ status",
       color = "Significance level (FDR corrected)") +
  th +
  geom_vline(xintercept = 0, linetype = "dotted", colour = "grey3") +
  scale_shape_manual(values = c("Negative" = 15, "Positive" = 16))+
  scale_color_manual(values = c("trend" = "grey3", "significant" = "red", "normal" = "grey48"))+
  guides(colour = "none")
  


ggsave("G:/My Drive/ADPD 2026/plot_all_regressions_estimates.png", width = 8, height = 6)
```


#Table

```{r}
library(gtsummary)
library(gt)


data_table<-data%>%
  filter(sex==2)%>%
  drop_na(age_v2_mri)%>%
  drop_na(A)%>%
  dplyr::select(age_v1, kids_total, A, APOE_binary)%>%
  mutate(A=factor(A),
         APOE_binary=factor(APOE_binary))%>%
  rename( `Number of childbirths` = kids_total,
         `Aβ+` = A,
         `APOE-ε4 carrier` = APOE_binary,
         `Age at first visit` = age_v1)



desc_table<-tbl_summary(data_table,
            by = `Number of childbirths`, 
            missing = "no")%>%
  add_n()%>%
  add_p%>%
  modify_header(label = "**Variable**") %>% # update the column header
  bold_labels()


desc_table%>%
  as_gt()%>%
  gtsave(filename = "G:/My Drive/ADPD 2026/table.docx")
#modify the total number of participants depending on availability of T1 and PACC difference
```




#Abstract

Objectives: Little is known about the impact of pregnancy on Alzheimer's Disease (AD) trajectory. We examined the impact of childbirths and AD pathology on cognitive decline, and related it to hippocampal volume change in cognitively unimpaired (CU) post-menopausal women at risk of AD.

```{r, echo=F}

pt_number<-data%>%
  drop_na(age_v2_mri)%>%
    drop_na(A)%>%
  nrow()%>%
  as.numeric()
  
pt_number_women<-data%>%
  drop_na(age_v2_mri)%>%
    drop_na(A)%>%
  filter(sex==2)%>%
  nrow()%>%
  as.numeric()

```

Methods: Our sample includes `r pt_number_women` CU post-menopausal women from the ALFA+ study. Hippocampus subfield segmentation was obtained using Freesurfer7. Amyloid (Aβ) positivity was defined as CSF Aβ42/40 ratio\<0.071. Cognition was measured at visit 1 and 2 using PACC scores. We first examined the impact of childbirths on cognitive change, and moderation by Aβ positivity. We then assessed the effect of childbirth on hippocampal subfield volume change, and moderation by Aβ positivity. All linear models included APOE-ε4 carriership, age, time difference between visits and total intracranial volume as covariates (Table 1). FDR correction was used in models 3 and 4.

Results: A trending interaction between number of childbirths and amyloid positivity on cognitive change was found (F=`r tidy_mod1_es[8]`, p=`r tidy_mod1_p[8]`), where Aβ+ women showed worse decline with higher parity. Childbirths had a significant positive main effect on volume change in the fimbria (F=5.240, adj.p.=0.020). A significant interaction of childbirths and Aβ+ was found in the parasubiculum (F=6.8155, adj.p.=0.0443), such that Aβ+ women showed worse volume reduction with higher parity. This interaction was trending in the CA4, dentate gyrus, presubiculum and subiculum.

Conclusion: In CU post-menopausal women at risk of AD, parity might moderate the impact of AD pathology on cognitive change at 3 years. Additionally, parity independently influences volume change in the fimbria, and interacts with AD pathology on parasubiculum volume. Results highlight the effect of parity on AD-related neuropathology via hippocampal volume change. 

```{r}
stop("Script stopped here intentionally")
```


# Miscellaneous

## test for model estimates of cognition?
```{r}
data_forest_plot_cog<-mediation_data%>%
  filter(sex == 2 )

models<-list(
  Attention = lmer(Z_Attention_Composite ~ kids_total*A +age_cog + APOE_binary + education  + Fuma2 +(1 | subject), data_cog_long_composites[data_cog_long_composites$sex == 2,]),
  Memory = lmer(Z_Memory_Composite ~ kids_total*A +age_cog + APOE_binary + education + Fuma2 + (1 | subject), data_cog_long_composites[data_cog_long_composites$sex == 2,]),
  Executive = lmer(Z_Executive_Composite ~ kids_total*A +age_cog + APOE_binary + education + Fuma2 + (1 | subject), data_cog_long_composites[data_cog_long_composites$sex == 2,]),
  Language = lmer(Z_Language_Composite ~ kids_total*A +age_cog + APOE_binary + education + Fuma2 + (1 | subject), data_cog_long_composites[data_cog_long_composites$sex == 2,]),
  Visual = lmer(Z_Visual_Composite ~ kids_total*A +age_cog + APOE_binary + education + Fuma2 + (1 | subject), data_cog_long_composites[data_cog_long_composites$sex == 2,])
)

# Extract coefficients and CIs
#install.packages("broom.mixed")
#library(broom.mixed)
#library(purrr)

pred_cog_all <- map_dfr(models, ~ {
  broom.mixed::tidy(.x, terms = c("kids_total", "A"))   
}, .id = "cog")%>%
  filter(term %in% c("kids_total", "A", "kids_total:A"))%>%
  dplyr::select(cog, term, estimate, std.error)%>%
  pivot_wider(names_from = term, values_from = c(estimate, std.error))%>%
  mutate(plot_estimate_kids_A0 = estimate_kids_total,
         plot_estimate_kids_A1 = estimate_kids_total + `estimate_kids_total:A`,
         plot_conf.low_kids_A0 = estimate_kids_total - 1.96 * std.error_kids_total,
         plot_conf.high_kids_A0 = estimate_kids_total + 1.96 * std.error_kids_total,
         plot_conf.low_kids_A1 = (estimate_kids_total + `estimate_kids_total:A` - 1.96 * `std.error_kids_total:A`),
         plot_conf.high_kids_A1 = (estimate_kids_total + `estimate_kids_total:A` + 1.96 * `std.error_kids_total:A`))%>%
  dplyr::select(-estimate_kids_total, -estimate_A,-`estimate_kids_total:A` ,-`std.error_kids_total`, -std.error_A ,         -`std.error_kids_total:A`)%>%
  pivot_longer(
    cols = starts_with("plot_"),  # Keep 'term' as-is
    names_to = c(".value", "A"),  # .value creates columns like estimate, conf.low, etc.
    names_pattern = "plot_(.*)_kids_A(\\d+)"
  )%>%
  mutate(colour_group = case_when(region == "CA4" ~ "trend",
                                  region == "GC-ML-DG" ~ "trend",
                                  region == "Presubiculum" ~ "trend",
                                  region == "Subiculum" ~ "trend",
                                  region == "Parasubiculum" ~ "significant",
                                  .default = "normal"))



pred_cog_all %>%
  mutate(A = factor(A, levels = c("1", "0"), labels = c("Positive", "Negative"))) %>%
  ggplot(aes(x = estimate, y = cog, group = A, colour = A)) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high), 
                  position = position_dodge(width = 0.8),
                  size = 0.6) +
  labs(x = "Estimated effect of parity on cog trajectory", 
       y = "Cog domain") +
  scale_shape_manual(values = c("Negative" = 15, "Positive" = 16))
  


#ggsave("G:/My Drive/ADPD 2026/plot_all_regressions_estimates.png", width = 8, height = 6)
```

### plotting both timepoints

uglyyy
```{r}
hippo_data_long%>%
  filter(sex== 2)%>%
  drop_na(A)%>%
  mutate(A=factor(A))%>%
  ggplot(aes(x=Visit, y=vol_parasubiculum, color = A))+
  geom_violin(size=1.1, position = position_dodge(width = 0.5))+
  geom_line( aes(group = subject), alpha=.2, size=1, position = position_dodge(width = 0.5))+
  geom_point( aes(group = subject), alpha=.2, size=1.3, position = position_dodge(width = 0.5))+
  facet_wrap(~kids_total)
```

### plotting differences
uglyyyy
```{r}
#I'm aware how clunky this code is... I was fighting with pivots for a while and getting nowhere. So this is the best I could do for now sorry! 

hippo_data_plot<-hippo
colnames(hippo_data_plot) <- paste(colnames(hippo_data_plot), 'V1', sep = '_')
hippo_data_plot<-hippo_data_plot%>%
  rename(subject = subject_V1)%>%
  dplyr::select(-Visit_V1)

hippo_data_plot2<-hippo2
colnames(hippo_data_plot2) <- paste(colnames(hippo_data_plot2), 'V2', sep = '_')
hippo_data_plot2<-hippo_data_plot2%>%
  rename(subject = subject_V2)%>%
  dplyr::select(-Visit_V2)

hippo_data_plot<-left_join(hippo_data_plot, hippo_data_plot2)

base_names <- names(hippo_data_plot) %>%
  str_subset("^vol_.*_V1$") %>%
  str_remove("_V1$")


for (base in base_names) {
  v1_col <- paste0(base, "_V1")
  v2_col <- paste0(base, "_V2")
  diff_col <- paste0(base, "_diff")
  
  hippo_data_plot[[diff_col]] <- hippo_data_plot[[v2_col]] - hippo_data_plot[[v1_col]]
}

hippo_data_plot<-hippo_data_plot%>%
  dplyr::select(subject, vol_CA1_diff, vol_CA3_diff, vol_CA4_diff, vol_GC_ML_DG_diff, vol_HATA_diff, vol_Hippocampal_tail_diff, vol_Whole_hippocampal_body_diff, vol_Whole_hippocampal_head_diff, vol_Whole_hippocampus_diff, vol_fimbria_diff, vol_molecular_layer_HP_diff, vol_parasubiculum_diff, vol_presubiculum_diff, vol_subiculum_diff)


hippo_data_plot_all <- left_join(data, hippo_data_plot) %>%
  mutate(across(starts_with("vol_"), ~ .x / diff_mri, .names = "{.col}_annual"))


hippo_data_plot_all%>%
  filter(sex==2)%>%
  mutate(A=factor(A))%>%
  drop_na(A)%>%
ggplot(aes(x=kids_total, y=vol_parasubiculum_diff_annual, group=A, colour=A))+
  geom_jitter()+
  geom_smooth(method = "lm")+
  call_aesthethics(15)+
  labs(x = "Number of childbirths", y = "Volume change")
```

## PACC ~ childbirth*AB - not in use anymore because using mixed effect model

```{r}

mod_pacc_w<-lm(PACC_diff ~ kids_total*A + A+age_v1 + education + APOE_binary + time_diff, data = data_cog_diff[data_cog_diff$sex==2,])
pred_pacc<-ggpredict(mod_pacc_w, terms=c("kids_total", "A"))


data_cog_diff%>%
  filter(sex == 2)%>%
  drop_na(A)%>%
  ggplot(aes(x=kids_total, y=PACC_diff, group=A, colour=A))+
  geom_jitter(width = 0.1)+
  geom_line(data=pred_pacc, 
            aes( x=x,
                 y=predicted,
                 group=group,
                 colour=group), 
            inherit.aes = F, 
            size=1)+
  geom_ribbon(data = pred_pacc, 
              aes(x=x, 
                  ymin= conf.low, 
                  ymax = conf.high, 
                  group=group), 
              inherit.aes = F, 
              alpha=.2)+
  th+
  labs(x = "Number of childbirths", y = "PACC change (Z-score)", group="Aβ+", colour="Aβ+")+
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey0")+
  scale_color_manual(values = c("0" = "steelblue2", "1" = "tomato"),
                     labels = c("0" = "Negative", "1" = "Positive"))+
  scale_fill_manual(values = c("0" = "steelblue2", "1" = "tomato"),
                    labels = c("0" = "Negative", "1" = "Positive"))


ggsave("G:/My Drive/ADPD 2026/plot_PACC_interaction.png", width = 8, height = 5)
```

##Cognitive results - LM of change
test
test again
```{r}
composites1<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 1 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -Z_PACC_Composite_1.2._v1 )%>%
  rename(subject = IdParticipante)

composites2<-read_excel("G:/My Drive/Data/ALFA+/Cognition/Visit 2 - Cognition Z Scores.xlsx")%>%
  dplyr::select(-Gender, -Age_V1S1, -Years_Edu, -V1S1_dt, -V2S1_dt, -Z_PACC_Composite_1.2._V2 )%>%
  rename(subject = IdParticipante)

PACC_diff<-read_excel("G:/My Drive/Data/ALFA+/Cognition/PACCv1_to_v3(incomplete).xlsx")%>%
  dplyr::select(IdParticipante, Visit, Z_PACC, FechaResultado)%>%
  rename(subject = IdParticipante)%>%
  mutate(subject = as.numeric(subject))%>%
  filter(Visit != 3)%>%
  pivot_wider(names_from = Visit, values_from = c(Z_PACC, FechaResultado))%>%
  mutate(PACC_diff = Z_PACC_2 - Z_PACC_1,
         time_diff = as.numeric(difftime(FechaResultado_2, FechaResultado_1, units= c("days"))))%>%
  dplyr::select(subject, PACC_diff, time_diff)

composite_all_diff<-full_join(composites1, composites2)%>%
  mutate(diff_Attention = Z_Attention_Composite_V2 - Z_Attention_Composite_v1,
         diff_Memory = Z_Memory_Composite_V2 - Z_Memory_Composite_v1,
         diff_Executive = Z_Executive_Composite_V2 - Z_Executive_Composite_v1,
         diff_Language = Z_Language_Composite_V2 - Z_Language_Composite_v1,
         diff_Visual = Z_Visual_Composite_V2 - Z_Visual_Composite_v1,
         subject = as.numeric(subject))%>%
  dplyr:: select(subject, diff_Attention, diff_Memory, diff_Executive, diff_Language, diff_Visual)

data_cog_diff<-left_join(data, composite_all_diff)%>%
  left_join(., PACC_diff)%>%
  mutate(A=factor(A),
         APOE_binary=factor(APOE_binary),
         AT=factor(AT))

mod_pacc_w<-lm(PACC_diff ~ kids_total*A +age_v1 + education + APOE_binary + time_diff, data = data_cog_diff[data_cog_diff$sex==2,])
summary(mod_pacc_w)

tidy_mod1_es<-tidy(mod_pacc_w)$estimate
tidy_mod1_p<-tidy(mod_pacc_w)$p.value


hist(data_cog_diff$PACC_diff)

plot_model(mod_pacc_w, type = "int")

#trying with poly
mod_pacc_w_poly<-lm(PACC_diff ~ poly(kids_total, 2, raw = TRUE)*A+ age_menopause_pmm +age_v1 + education + APOE_binary + time_diff, data = data_cog_diff[data_cog_diff$sex==2,])
summary(mod_pacc_w_poly)
```


## test multivariate MEM
multivariate MEM test
interesting time interaction with ptau217
```{r}

test_longer<- data_cog_long_composites%>%
  dplyr::select(-Z_aWMS_DR_RS, -Z_FCSRTI_RT, -Z_FS_totalcnswr, -Z_Wclave_totalscr)%>%
  pivot_longer(cols = starts_with("Z_"), names_to = "cog_domain", values_to = "cog_score")%>%
  mutate(cog_domain = factor(cog_domain))

fit_cog_domain_test<-lmer(cog_score ~ kids_total*A +age_cog + cog_domain + APOE_binary + education + (1 | subject), test_longer[test_longer$sex == 2,])
summary(fit_cog_domain_test)

fit_cog_domain_test<-lmer(cog_score ~ kids_total*CSF_pTau217*time_diff +age_v1 + cog_domain + APOE_binary + education + (1 | subject), test_longer[test_longer$sex == 2,])
summary(fit_cog_domain_test)


plot_model(fit_cog_domain_test, type= "int")


```
